### 2018-02-01 article/article.tex
add latex version

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Blue Hunters: Bluetooth RSSI Locator Robots},
            pdfauthor={Jacob Glueck (jng55); Jane Du (zd53); Justin Cray (jgc232)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\graphicspath{{../assets/}{../renderings/}}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Blue Hunters: Bluetooth RSSI Locator Robots}
\providecommand{\subtitle}[1]{}
\subtitle{\emph{ECE 4760, Fall 2017}}
\author{Jacob Glueck (\href{mailto:jng55@cornell.edu}{jng55}) \and Jane Du (\href{mailto:zd53@cornell.edu}{zd53}) \and Justin Cray (\href{mailto:jgc232@cornell.edu}{jgc232})}
\date{December 6, 2017}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

We built 2 small cars which used Bluetooth Received Signal Strength
Indicator (RSSI) measurements to navigate towards a stationary base
station. The cars and base station used a Bluetooth Low Energy (BLE) 4.0
module to take the measurements and a PIC32MX250 microcontroller. The
cars also used a 3 axis magnetometer as as compass in order to reliably
turn, as well as 2 micro 9 g servos to drive. Each unit was powered with
3 AA batteries. Finally, the chassis and wheels of each car were 3D
printed.

\hypertarget{video}{%
\subsubsection{Video}\label{video}}

\hypertarget{design}{%
\subsection{Design}\label{design}}

\hypertarget{chassis}{%
\subsubsection{Chassis}\label{chassis}}

The robots are made from 4 3D printed pieces: 2 wheels, the frame, and
the caster in the back. The servos, a 3 AA battery holder, and a
perfboard containing all the circuitry are mounted directly to the
frame.

The robots where designed in \href{http://www.openscad.org/}{OpenSCAD},
and their source code is available in
\href{https://github.com/orangeturtle739/bluehunters/tree/master/cad}{our
git repository} and
\protect\hyperlink{appendix-b-source-listing}{below}. There are three
files, \href{generated/frame.scad.html}{\texttt{frame.scad}},
\href{generated/drag.scad.html}{\texttt{drag.scad}}, and
\href{generated/wheel.scad.html}{\texttt{wheel.scad}}, for each of the
three parts. The main modules are defined in
\href{generated/main.scad.html}{\texttt{main.scad}}, and the other files
just instantiate them. The following renderings show each part:

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{frame.png}
\caption{Robot chassis}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{drag.png}
\caption{Caster wheel on back of robot}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{wheel.png}
\caption{Robot wheel}
\end{figure}

The parts were printed in ABS using
\href{https://www.monoprice.com/product?c_id=107\&cp_id=10724\&cs_id=1072403\&p_id=13860\&seq=1\&format=2}{Maker
Select 3D Printer v2} printers. All parts were printed with a layer
height of 0.3 mm, as there was no need for a smooth finish or high
tolerances. The parts where sliced with
\href{https://ultimaker.com/en/products/ultimaker-cura-software}{Cura}.

\hypertarget{pictures}{%
\subsubsection{Pictures}\label{pictures}}

Full system (base station and 2 robots):

\includegraphics[width=0.5\textwidth,height=\textheight]{full_system.jpg}

Multiple views of the robots:

\begin{longtable}[]{@{}ll@{}}
\toprule
\includegraphics{top_view.jpg} &
\includegraphics{side_view.jpg}\tabularnewline
\includegraphics{front_view.jpg} &
\includegraphics{inside_view.jpg}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{electronics}{%
\subsubsection{Electronics}\label{electronics}}

\hypertarget{bluetooth-modules}{%
\paragraph{Bluetooth modules}\label{bluetooth-modules}}

The HM-10 bluetooth modules we bought off Ebay were fakes: they were not
made by Jnhuamao \footnote{Jnhuamao is the company which makes the HM-10
  module, along with other Bluetooth modules. Their English website is:
  \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}, and did not come
with genuine Jnhuamao firmware. However, they did have real TI CC2541
chips. We realized they were fakes when they did not behave according to
the Jnhuamao data sheet (see \protect\hyperlink{data-sheets}{the data
sheets section}). Luckily, the hardware on the fake chips is the same as
that of the genuine chips, minus an external crystal, and the genuine
firmware checks for the presence of the crystal, and works even without
it. \footnote{\url{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}.}
As such, we reprogrammed the chips with the genuine firmware according
to an
\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
forum post}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We soldered wires to the programming pins on the breakout boards, and
  connected those pins to an
  \href{https://www.pjrc.com/teensy/teensy31.html}{Arduino Teensy 3.2}.
  We chose a Teensy because it is 3.3 V as opposed to 5, which would
  damage the 3.3 V CC2541.

  The pins were connected as follows:

  \begin{longtable}[]{@{}lll@{}}
  \toprule
  Name & CC2541 Pin & Arduino Pin\tabularnewline
  \midrule
  \endhead
  \texttt{DEBUG\_CLOCK} & 7 & 5\tabularnewline
  \texttt{DEBUG\_DATA} & 8 & 6\tabularnewline
  \texttt{RESET\_N} & 11 & 4\tabularnewline
  \bottomrule
  \end{longtable}

  The layout of the HM-10 board is:

  \includegraphics{hm10_pins.png}
\item
  We uploaded the
  \href{https://github.com/RedBearLab/CCLoader/blob/master/Arduino/CCLoader/CCLoader.ino}{\texttt{CCLoader.ino}
  sketch} to the Arduino.
\item
  Finally, we ran (in a Windows virtual machine)
  \href{https://github.com/RedBearLab/CCLoader/tree/master/Windows}{\texttt{CCLoader.exe}}.
  This program takes 3 arguments:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{CCLoader.exe} \OperatorTok{<}\NormalTok{COM Port}\OperatorTok{>} \OperatorTok{<}\NormalTok{Firmware.bin}\OperatorTok{>}\NormalTok{ 0}
\end{Highlighting}
\end{Shaded}

  The firmware file came from the same Arduino form post, and can be
  found
  \href{http://forum.arduino.cc/index.php?action=dlattach;topic=393655.0;attach=183702}{here}.
\end{enumerate}

There is also an excellent
\href{https://www.youtube.com/watch?v=ez3491-v8Og}{YouTube video}, which
explains the firmware flashing process.

After flashing genuine firmware onto the chip, the next step was to
update the firmware to the latest version. The firmware flashed onto the
board was version 540, but Jnhuamao had (at the time we did this
project)
\href{http://www.jnhuamao.cn/rom/HMSoft-10-2541-V603.zip}{released
version 603}. \footnote{Jnhuamao's firmware download page
  \url{http://www.jnhuamao.cn/download_rom_en.asp?id=1\#}.} They also
provide
\href{http://www.jnhuamao.cn/HowToUpgradeFirmware_en.zip}{instructions
on how to upgrade the firmware}. The basic process is:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Connect the HM-10 module to a computer using a 3.3 V FTDI to USB
  adapter. Then, use PUTTY to establish a serial session (9600 baud,
  8N1). Send the chip \texttt{AT}; if it is connected properly, it will
  respond with \texttt{OK}.
\item
  Send the chip \texttt{AT+SBLUP} to put it in firmware update mode. It
  will respond with \texttt{OK+SBLUP}. Terminate the PUTTY session.
\item
  Run the \texttt{HMSoft.exe} program distributed in the firmware update
  download. We ran it in a Windows virtual machine because we did not
  trust it.
\item
  Select the proper port and firmware file using the software, and hit
  ``Load Image''. The software should handle the rest!
\item
  To make sure it worked, establish a serial connection again using
  PUTTY. Send \texttt{AT+VERS?} to query the chip for version
  information.
\end{enumerate}

\hypertarget{software}{%
\subsubsection{Software}\label{software}}

The robots were programmed in C, using the
\href{http://www.microchip.com/mplab/mplab-x-ide}{MPLAB X IDE} v4.0,
with the \href{http://www.microchip.com/mplab/compilers}{XC32} v1.4
compiler and the
\href{http://www.microchip.com/SWLibraryWeb/product.aspx?product=PIC32\%20Peripheral\%20Library}{PIC
32 Legacy Peripheral Library (plib)}. The full source code is available
in
\href{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}{our
git repository} and
\protect\hyperlink{appendix-b-source-listing}{below}. The code is
divided into 5 main units:

\begin{itemize}
\tightlist
\item
  \href{generated/ble.c.html}{\texttt{ble.c}}: contains all the
  functions for interacting with the BLE device over UART.
\item
  \href{generated/imu.c.html}{\texttt{imu.c}}: contains all the
  functions for interacting with the IMU (which contained the
  magnetometer) over I2C.
\item
  \href{generated/servo.c.html}{\texttt{servo.c}}: contains all the
  functions for interacting with the servos using PWM.
\item
  \href{generated/main.c.html}{\texttt{main.c}}: the main code.
\item
  \href{generated/pt_cornell_1_2_2.c.html}{\texttt{pt\_cornell\_1\_2\_2.c}}:
  contains the functions which were originally declared in the
  protothreads header file. However, since they were in the header file,
  if multiple source files included the header file, there would be
  linking errors due to duplicate definitions of symbols. Moving the
  protothreads definitions to a separate file resolved this issue.
\end{itemize}

\hypertarget{ble}{%
\paragraph{BLE}\label{ble}}

The BLE module used UART, at 9600 baud 8N1. We used \texttt{UART1} on
the PIC for communicating with the BLE module, and used \texttt{UART2}
for communicating with the computer (for debugging). The BLE files
define a series of macros and \texttt{PT\_THREAD} functions for printing
characters and reading lines in a non-blocking fashion using
protothreads. All of the \texttt{PT\_THREAD} functions can be spawned
with \texttt{PT\_SPAWN} to run them.

The \href{generated/main.c.html}{\texttt{main.c}} file contains a thread
which does all the communication with the Bluetooth chip using these
functions. The commands it sends to the chip are:

\begin{itemize}
\item
  \texttt{AT+RESET}: resets the chip to ensure it is in a clean state
  before receiving other commands.
\item
  \texttt{AT+IBEA1}: enables the iBeacon functionality of the chip (sets
  it to 1; \texttt{AT+IBEA0} would disable it by setting it to 0). This
  allows the chip to be found with an RSSI scan. After setting the
  value, we query it with \texttt{AT+IBEA?} and send the result over the
  other serial port to a computer for debugging.
\item
  \texttt{AT+ROLE0} or \texttt{AT+ROLE1}: sets the role to either
  peripheral (0) or central (1). The base station is set to peripheral,
  and the 2 robots are set to central. Peripheral means the device will
  respond in inquiries from a central device. This allows it to be
  discovered during an RSSI scan. After setting the value, we query it
  with \texttt{AT+ROLE?} and send the result over the other serial port
  to a computer for debugging.
\item
  \texttt{AT+IMME0} or \texttt{AT+IMME1}: sets the work state of the
  device to either actively listening for Bluetooth signals (0), or only
  acting when it receives a serial command (1). Once again, the base
  station is set to 0: it needs to listen for signals and respond. The
  robots are set to 1, as the chips need to initiate scan requests when
  they receive the command over serial. After setting the value, we
  query it with \texttt{AT+IMME?} and send the result over the other
  serial port to a computer for debugging.
\item
  \texttt{AT+NAME\%s}: sets the name of the chip (which is visible when
  scanning) to \texttt{\%s} (For example, \texttt{AT+NAMEPIRATE} names
  the chip \texttt{PIRATE}). We give all the chips unique names to make
  debugging easier. After setting the value, we query it with
  \texttt{AT+NAME?} and send the result over the other serial port to a
  computer for debugging.
\item
  \texttt{AT+SHOW3}: configures the device to advertise both its name
  and RSSI when scanning.
\item
  \texttt{AT+ADDR?}: queries the device for its hardware address. We
  recorded the hardware device of each chip, as when doing RSSI scans,
  the results are reported by hardware address.
\item
  \texttt{AT+DISI?}: performed only on the robots, causing a discovery
  scan. The result of the scan is a bunch of lines of the form:

\begin{verbatim}
OK+DISC:00000000:00000000000000000000000000000000:0000000000:6832A3801EBE:-080
\end{verbatim}

  The second to last token, \texttt{6832A3801EBE}, is the hardware
  address of the discovered device, and the last token, \texttt{-080},
  is the measured RSSI. The chip will transmit a line for each device it
  finds (``line'' is a misnomer as it does not separate them with any
  characters), followed by \texttt{OK+DISCE}.
\end{itemize}

One interesting thing to note about the chip is that commands do not
have to end with newlines or carriage returns. However, if sent, the
chip will ignore them.

\hypertarget{imu}{%
\paragraph{IMU}\label{imu}}

The PIC commmunicates with the IMU via I2C. The IMU includes a breakout
board for the QFN MPU-9250 module, which itself includes 2 dies. One
contains the 3-axis gyroscope and 3-axis accelerometer, which were not
used in this project, and the other die is the AK8963 3-axis
magnetometer (compass).

It is connected to the rest of the MPU module via an auxillary I2C bus,
so it is not connected to the MPU's main I2C bus by default. While the
accelerometer and gyroscope registers can be read after powering up the
IMU, the compass also needs pass-through mode to be enabled on the IMU
to make it an accessible slave on the I2C bus.

Some useful functions as defined in
\href{generated/imu.h.html}{\texttt{imu.h}} are described below:

\begin{itemize}
\tightlist
\item
  \texttt{void\ imu\_init()}: Initializes the MPU-9250, including
  configuring the chip to allow reading the compass.
\item
  \texttt{int\ imu\_get\_heading()} Returns the heading of the robot as
  a value between -180 and 180. The compasses were not completely
  calibrated to find magnetic north; it only ensures angles are correct
  relevant to past headings.
\item
  \texttt{void\ imu\_mag\_read\_data(int\ *\ destination)} Fetches
  compass readings; saves register values into \texttt{destination} in
  the form \texttt{{[}x,\ y,\ z{]}}.
\item
  \texttt{int\ angle\_diff(int\ source,\ int\ target)} Gets the
  difference between two angles in degrees to account for discontinuity
  between -180 and 180 degrees.
\item
  \texttt{int\ degree(int\ deg)}: Offsets degree values so that they
  fall within the range -180 to 180 degrees.
\end{itemize}

Initializing the IMU involves opening the I2C module, and then
configuring the IMU.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We open the \texttt{I2C2} module; \texttt{I2C1} uses pins already used
  by the connections to the Bluetooth module. We open it with the baud
  rate generator value
  \texttt{BRG\ =\ (Fpb\ /\ 2\ /\ baudrate)\ -\ 2\ =\ 4e7\ /\ 2\ /\ 4e5\ -\ 2\ =\ 48},
  as specified for \texttt{OpenI2C2()} in the
  \protect\hyperlink{references}{peripheral libraries}.
\item
  Pass through is enabled, interrupts for data ready are enabled, the
  IMU as an I2C master function is disabled, and the sensor is powered
  up.
\end{enumerate}

This enables the PIC to talk to the AK8963. The AK8963 has several modes
of operation, and the chip must be set to power-down mode before
switching to other modes. We read compass values with single measurement
mode, as specified below:

\includegraphics{imu_single_measurement.png}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Set the compass to single measurement mode in 14 bit resolution.
\item
  Read the 6 data registers (X low, X high, Y low, Y high, Z low, Z
  high)
\item
  Read the Status 2 register to check for magnetic sensor overflow.
  Without reading this register, the read is not considered complete and
  further reads will fail.
\item
  Wait; if the IMU is read too frequently, it will not have enough time
  to take measurements.
\end{enumerate}

Additional helper methods used in I2C were defined in
\href{generated/imu.c.html}{\texttt{imu.c}}. These are largely based on
the files from the
\protect\hyperlink{code-and-designs-borrowed-from-others}{self-balancing
robot}.

\begin{itemize}
\tightlist
\item
  \texttt{char\ i2c\_read\_device(char\ device,\ char\ address)} Reads
  the data from a single register at \texttt{address}
\item
  \texttt{void\ i2c\_write\_byte(char\ device,\ char\ address,\ char\ data)}
  All configurations used in this project involved writing single bytes
  of data.
\item
  \texttt{void\ i2c\_wait(int\ cnt)} Writes 2 nops; reads require time
  to return a value, and calling reads consecutively.
\end{itemize}

In order to calibrate the compass, the robots spun in place when powered
on. They recorded the maximum and minimum values for each axes, and used
that data to scale and center the magnetometer readings.

\hypertarget{servos}{%
\paragraph{Servos}\label{servos}}

For the drive systems, we used the FS90R servos. These servos are
continuous rotation servos with a stall torque of 1.5 kg-cm. Continuous
rotation servos operate based on PWM duty cycle; since servos operate at
a pulse width of 1 to 2 ms out of 20 ms, we used a timer with a
prescaler of 32 and a maximum count of 25600, and set the pwm at numbers
ranging from 1280 to 2560 (corresponding to the right pulse width).

A 50\% duty cycle means the servos are still. A less than 50\% duty
cycle means the servo turns clockwise and a greater than 50\% duty cycle
means the servo turns counter-clockwise. To implement the PWM, we use
the output capture on the PIC. We set the duty cycle using the built in
timers. We set a max and min duty cycle value which were experimentally
derived. The middle point is the average of the two. Tank drive is used
to control the system where both wheels are driven forward or point
turns where one wheel is turned in one direction and one is turned in
the other direction.

\hypertarget{protothreads}{%
\paragraph{Protothreads}\label{protothreads}}

While the provided protothreads library was very helpful, it did not
work correctly when used in multiple compilation units. The reason for
this was that it defined variables and functions in the header file, so
if multiple source files included
\href{generated/pt_cornell_1_2_2.h.html}{\texttt{pt\_cornell\_1\_2\_2.h}},
linking would fail with duplicate symbol definitions. Furthermore, if
multiple files included
\href{generated/config_1_2_2.h.html}{\texttt{config\_1\_2\_2.h}},
linking would fail with the obscure error:

\begin{verbatim}
section `.config_BFC00BF4' will not fit in region `config2'
\end{verbatim}

This happened because if multiple compilation units included
\texttt{\#pragma\ config} directives (which they did by including a
\texttt{\#include\ "config\_1\_2\_2.h"} directive), then the linker
would try to assign too many symbols to the configuration region.

The fix to this was twofold:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We modified protothreads to have both a header and source file, with
  definitions only in the source file.
\item
  We only included
  \href{generated/config_1_2_2.h.html}{\texttt{config\_1\_2\_2.h}} in
  the main C file, and moved all macro definitions to the protothreads
  header.
\end{enumerate}

The updated versions of protothreads can be found in
\protect\hyperlink{appendix-b-source-listing}{Appendix B}.

\hypertarget{gradient-descent}{%
\paragraph{Gradient descent}\label{gradient-descent}}

The algorithm for deciding what path to follow is a basic version of
gradient descent. The following image represents the decision-making
state machine, where the starting state is \textbf{Measure rssi twice,
take average}.

\includegraphics{grad_desc.png}

We also implemented and tested the following improved version that
allows for correction; a car that has just moved forward and detected a
weakened signal does not know whether the beacon to its left or right.
If after turning, the signal is still weaker, it has picked the wrong
turn. This decision process corrects this (same starting state:
\textbf{Measure rssi twice, take average}):

\includegraphics{turn_correction.png}

It did not prove much more accurate than randomized gradient descent,
largely due to noisy readings from IMU and Bluetooth signal strength. As
such, we used the simplified version.

\hypertarget{testing}{%
\subsubsection{Testing}\label{testing}}

\hypertarget{ble-1}{%
\paragraph{BLE}\label{ble-1}}

Once we flashed the Bluetooth chips with the proper firmware, we used
the
\href{https://itunes.apple.com/us/app/lightblue-explorer/id557428110?mt=8}{LightBlue}
app to test communications. Using the app, we could discover the
Bluetooth chips, and send and receive data. By initially using a the
iPhone app, which was known to work, we could debug the Bluetooth chips
in isolation. Once that worked, we had the chips communicate with each
other. Finally, we had the chips take RSSI measurements.

We also experimented with RSSI, as described in
\protect\hyperlink{results}{results}. This data validated the
effectiveness of RSSI as an approximate metric for distance measurement,
though confirmed that it would be highly variable.

\hypertarget{imu-1}{%
\paragraph{IMU}\label{imu-1}}

\begin{itemize}
\tightlist
\item
  Basic I2C: Basic testing was required to ensure I2C communication with
  the chip was working. This included testing the correctness of the I2C
  read and write functions (including the order of starts, restarts, and
  idles in accordance with the I2C protocol and the chip datasheets), as
  well basic parameters like the address of the MPU-9250 and its
  registers.
\item
  Compass I2C: This was more complex, and mostly involved testing the
  usage of the modes of both the MPU-9250 (IMU module) and the AK8963
  (the compass itself). Performance testing was also important here, as
  the compass would not allow rapid reads in quick succession; we
  inserted nops in the
  \texttt{imu\_mag\_read\_data(int\ *\ destination)} for this reason.
\item
  Compass calibration: We tested incrementally, first using hard-coded
  maximum and minimum observed values outputted from the IMU, and then
  with values taken from a self-calibration in which the robot would
  spin in a circle, and the PIC would record the largest and smallest
  values it saw on the X- and Y-axes.
\end{itemize}

\hypertarget{servos-1}{%
\paragraph{Servos}\label{servos-1}}

We tested the functionality of servo code by testing individual
functionality on both the hardware and software side. We tested the
servo code in stages, first implementing control at a fixed speed and
then implementing directional motion and lastly turning. This allowed us
to verify our parts of the code and the functionality of the servos.

\hypertarget{whole-system}{%
\paragraph{Whole system}\label{whole-system}}

\begin{itemize}
\tightlist
\item
  Some subsystem testing was required for the servo with the IMUs, for
  turning/driving straight with feedback. We debugged by reading in IMU
  values via UART and observing the robot's turns.
\item
  We tested extensively with the beacon and two hunters in different
  environments. Most of our initial testing was in a large, open space
  with a few metal tables on the periphery. We later tested in the lab
  in the Digital Lab (Phillips 238), and in the hallway where our demo
  took place. The confined space of the inside of the lab and hallway,
  as well as the presence of more reflective bodies (metal doors,
  tables, people) negatively impacted the performance of our hunters,
  and it was necessary to recalibrate the hunters to work better in the
  demo environment.
\end{itemize}

\hypertarget{results}{%
\subsection{Results}\label{results}}

In most cases, at least 1 of the 2 robots successfully made it to the
base station. However, it was not as reliable as we initially hoped. One
of the main reasons for this was the noise in RSSI measurements.

We expected that RSSI would vary with distance according to the
following relation:

\(\text{RSSI} = A - 10 n \log(d)\)

where \(A\) and \(n\) are RF propagation parameters in dBm, \(d\) is
distance in meters, and RSSI is the measured RSSI in dBm. \footnote{L.
  Peneda, A. Azenha and A. Carvalho, ``Trilateration for indoors
  positioning within the framework of wireless communications,'' 2009
  35th Annual Conference of IEEE Industrial Electronics, Porto, 2009,
  pp.~2732-2737.

  doi: 10.1109/IECON.2009.5415423

  URL:
  \url{http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=5415423\&isnumber=5414636}}
We experimented with RSSI measurements to determine how well they worked
by taking 2 Bluetooth modules, and measuring the RSSI while changing the
distance between them. One remained stationary on the floor and the
other was moved away from it 1 floor tile (each floor tile is a 1 foot
square) at a time. At each point, we took 3 RSSI measurements and
averaged them. The graph below displays the results:

\includegraphics{rssi-chart.png}

While the chip reported RSSI in units proportional to dBm, and we
measured distances in feet, not meters, we could still use the above
formula without worrying about unit conversions. The constants \(A\) and
\(n\), provided we determined them empirically based on the data, would
encode the conversions. As such, we fit the data using the above formula
with \(A=-48\) and \(n=3\), resulting in the blue curve above. While the
general shape of the curve matches, there is significant noise in the
averaged RSSI data. Furthermore, when we tried to reproduce the
measurements, we could not do so precisely -- it seemed to even depend
on where our feet where! See \protect\hyperlink{rssi-data}{Appendix B}
for the source data for the table.

Despite how noisy the RSSI measurements where, the robots were still
able to perform a reasonably accurate gradient descent. In most cases,
at least one of the 2 robots would find the beacon in a matter of
minutes.

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

This assignment was an interesting exploration into short-range distance
determination using Bluetooth, a generally unconventional approach. We
were warned beforehand about the difficulties in using Bluetooth RSSI, a
measurement particularly susceptible to multipath interference, as a
distance sensor -- both in the papers and reference materials researched
before implementation, as well as by course staff in the early stages of
our project. We worked with the course staff to simplify and stratify
our project into workable milestones. Our final project is a reduced
version of our original plan.

Our hunting bots worked reliably when they stayed within roughly 1 m of
the beacon. After this, they entered the land of shallow gradients: the
signal strength from the beacon (already noisy) would not change very
much, and often only due to noise. The hunting bots normally could never
recover from this. Our bots also worked better in larger, more open
spaces, preferably without metal structures like tables or doors nearby.
Both of these are reasonable given the operating range of BLE 4.0.

We reused some of the code from the
\href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/index.html}{Self-Balancing
Robot} as
\protect\hyperlink{code-and-designs-borrowed-from-others}{cited} in our
references section, and explained in \protect\hyperlink{imu}{IMU}. The
rest of the design is ours, and is not an infringement on intellectual
property. We also used the Bluetooth firmware flashing technique we
discovered on the
\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
forums}.

We did not reverse-engineer any technology, and did not encounter any
issues with trademarks or patents. We did not have to sign
non-disclosure to get a sample part. We hope to further polish our
findings and report, and will aim to publish them in a print magazine,
ideally as a part of ECE 4920.

There are many potential avenues for improvements or further
development. In addition to more circumstantial difficulties encountered
during the project work period (such as extremely unbalanced servos,
where one always ran faster than the other), two areas of potential
improvement are as follows:

\begin{itemize}
\tightlist
\item
  \emph{Communication between the two robots}. While Bluetooth may not
  offer good distance measurement via RSSI, it can be used for reliable
  communication between modules. It would be straightforward for one
  hunting robot to inform the other whether it believes it is
  approaching the beacon or not. In the simplest case, a hunting robot
  that is approaching, or already at, the beacon can provide a second
  point of reference for a currently hunting robot.
\item
  \emph{More complete usage of IMU}. A lot of development time was spent
  simply on getting the PIC to compass communication running, and not
  much time was spent on calibrating the sensor or processing the data.
  In further stages of this project, we could continue to work towards
  dead reckoning of the mobile robots. This, paired with inter-swarm
  communication, would make for a much more sophisticated and likely
  much more efficient system. (Of course, this does not resolve the
  shallow gradients problem -- but it would allow the approach to the
  beacon to be much faster.)
\end{itemize}

\hypertarget{ieee-code-of-ethics}{%
\subsubsection{IEEE Code of Ethics}\label{ieee-code-of-ethics}}

Our project, as well as the design and creation project thereof, adheres
to the \href{https://www.ieee.org/about/ethics.html}{IEEE Code of
Ethics}. The project itself is relatively small and physically
self-contained; it did not have any immediate impact on the environment
or safety and health of potential users, we the creators, or any others.
We chose the project's topic out of interest and for a challenge,
without any conflict of interest. Our report is honest and realistic;
none of our data is fabricated, and our conclusions reflect our hard
findings. We did not accept any form of bribery.

The project investigated the use of an emerging and popular technology
(BLE) in a new application (distance sensing). We consistently received
and built upon advice from our TAs and professor Bruce Land, and cite
the work we used in the basis of our project.

Additionally, neither we nor our project presented any act of
discrimination, and our project does not injure others in any way. We
assisted our colleagues (both fellow group members and other classmates)
when the occasion arose, such as by offering help debugging or lending a
spare module before a demo.

Overall, our project complies with the IEEE Policies, 7.8 Code of
Ethics.

\hypertarget{appendices}{%
\subsection{Appendices}\label{appendices}}

\hypertarget{appendix-a-legal-considerations}{%
\subsubsection{Appendix A: Legal
Considerations}\label{appendix-a-legal-considerations}}

\begin{itemize}
\tightlist
\item
  The group approves this report for inclusion on the course website.
\item
  The group approves the video for inclusion on the course youtube
  channel.
\item
  The TI CC2541 meets FCC CFR47 Part 15 requirements. \footnote{See the
    top of page 8 of \url{http://www.ti.com/lit/ds/symlink/cc2541.pdf}
    for compliance information.}
\end{itemize}

\hypertarget{appendix-b-source-listing}{%
\subsubsection{Appendix B: Source
Listing}\label{appendix-b-source-listing}}

Our source code can be found
\href{https://github.com/orangeturtle739/bluehunters}{on Github}. Each
file can also be found below:

\hypertarget{code}{%
\paragraph{Code}\label{code}}

\begin{itemize}
\tightlist
\item
  \href{generated/ble.c.html}{\texttt{ble.c}}
\item
  \href{generated/ble.h.html}{\texttt{ble.h}}
\item
  \href{generated/imu.c.html}{\texttt{imu.c}}
\item
  \href{generated/imu.h.html}{\texttt{imu.h}}
\item
  \href{generated/servo.c.html}{\texttt{servo.c}}
\item
  \href{generated/servo.h.html}{\texttt{servo.h}}
\item
  \href{generated/pt_cornell_1_2_2.c.html}{\texttt{pt\_cornell\_1\_2\_2.c}}
\item
  \href{generated/pt_cornell_1_2_2.h.html}{\texttt{pt\_cornell\_1\_2\_2.h}}
\item
  \href{generated/main.c.html}{\texttt{main.c}}
\item
  \href{generated/main.h.html}{\texttt{main.h}}
\item
  \href{generated/config_1_2_2.h.html}{\texttt{config\_1\_2\_2.h}}
\end{itemize}

\hypertarget{cad}{%
\paragraph{CAD}\label{cad}}

\hypertarget{sources}{%
\subparagraph{Sources}\label{sources}}

\begin{itemize}
\tightlist
\item
  \href{generated/main.scad.html}{\texttt{main.scad}}
\item
  \href{generated/frame.scad.html}{\texttt{frame.scad}}
\item
  \href{generated/wheel.scad.html}{\texttt{wheel.scad}}
\item
  \href{generated/drag.scad.html}{\texttt{drag.scad}}
\item
  \href{generated/concept.scad.html}{\texttt{concept.scad}}
\end{itemize}

\hypertarget{stl}{%
\subparagraph{STL}\label{stl}}

\begin{itemize}
\tightlist
\item
  \href{stl/frame.stl}{\texttt{frame.stl}}
\item
  \href{stl/wheel.stl}{\texttt{wheel.stl}}
\item
  \href{stl/drag.stl}{\texttt{drag.stl}}
\end{itemize}

\hypertarget{rssi-data}{%
\paragraph{RSSI data}\label{rssi-data}}

\begin{longtable}[]{@{}llllll@{}}
\toprule
Distance (feet) & RSSI A & RSSI B & RSSI C & RSSI Average & Expected
RSSI\tabularnewline
\midrule
\endhead
1 & -48 & -48 & -48 & -48.00 & -48.00\tabularnewline
2 & -56 & -58 & -52 & -55.33 & -57.03\tabularnewline
3 & -66 & -66 & -68 & -66.67 & -62.31\tabularnewline
4 & -64 & -68 & -65 & -65.67 & -66.06\tabularnewline
5 & -60 & -70 & -64 & -64.67 & -68.97\tabularnewline
6 & -80 & -80 & -76 & -78.67 & -71.34\tabularnewline
7 & -80 & -79 & -65 & -74.67 & -73.35\tabularnewline
8 & -68 & -66 & -72 & -68.67 & -75.09\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{appendix-c-schematics}{%
\subsubsection{Appendix C: Schematics}\label{appendix-c-schematics}}

\includegraphics{schematic.png}

\hypertarget{appendix-d-bill-of-materials}{%
\subsubsection{Appendix D: Bill Of
Materials}\label{appendix-d-bill-of-materials}}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
\begin{minipage}[b]{0.15\columnwidth}\raggedright
Name\strut
\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright
Manufacturer Part Number\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Vendor\strut
\end{minipage} & \begin{minipage}[b]{0.17\columnwidth}\raggedright
Vendor Part Number\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
Quantity\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright
Unit Cost\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright
Total Cost\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.15\columnwidth}\raggedright
BLE 4.0 Module (TI CC2541)\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
HM-10\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\href{https://www.ebay.com/itm/AT-09-BLE-Bluetooth-4-0-Uart-Transceiver-Module-CC2541-Central-Switching-HM-10/142425748901?ssPageName=STRK\%3AMEBIDX\%3AIT\&_trksid=p2057872.m2749.l2649}{Ebay}\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
142425748901\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
3\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$3.99\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$11.97\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\raggedright
FEETECH FS90R (pack of 2) Continuous Rotation Robotic Servo\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
FS90R\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\href{https://www.amazon.com/gp/product/B074BFQC3Q/ref=oh_aui_detailpage_o06_s00?ie=UTF8\&psc=1}{Amazon}\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
B074BFQC3Q\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
2\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$12.39\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$24.78\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\raggedright
HiLetgo 9-Axis 9 DOF 16 Bit Gyroscope Acceleration Magnetic Sensor\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
MPU-9250\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\href{https://www.amazon.com/HiLetgo-Gyroscope-Acceleration-Accelerator-Magnetometer/dp/B01I1J0Z7Y/ref=redir_mobile_desktop?_encoding=UTF8\&dpID=51nl2fcMh6L\&dpPl=1\&keywords=mpu\%209250\&pi=AC_SX236_SY340_QL65\&qid=1512564044\&ref=plSrch\&ref_=mp_s_a_1_3\&sr=8-3}{Amazon}\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
B01I1J0Z7Y\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
2\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$8.49\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$16.98\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\raggedright
Small board\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
Lab rental\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
3\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$5.00\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$15.00\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\raggedright
PIC32MX250F128B\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
PIC32MX250F128B\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
Lab rental\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
3\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$5.00\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$15.00\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\raggedright
6-inch Protoboard\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
Lab rental\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
3\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$2.50\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$7.50\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.15\columnwidth}\raggedright
Male header pins\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
Lab\strut
\end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
--\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
129\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
\$0.05\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\$6.45\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\textbf{Total:} \$97.68

\hypertarget{appendix-e-work-distribution}{%
\subsubsection{Appendix E: Work
Distribution}\label{appendix-e-work-distribution}}

\begin{itemize}
\tightlist
\item
  \textbf{Jacob:}

  \begin{itemize}
  \tightlist
  \item
    Bluetooth firmware hacking
  \item
    Bluetooth UART code
  \item
    CAD design \& 3D printing
  \end{itemize}
\item
  \textbf{Jane:}

  \begin{itemize}
  \tightlist
  \item
    I2C IMU code
  \item
    Robot motion code and gradient descent algorithm
  \end{itemize}
\item
  \textbf{Justin:}

  \begin{itemize}
  \tightlist
  \item
    Servo control code
  \item
    Board assembly (3 small boards and our 3 of our protoboards)
  \item
    Schematic
  \end{itemize}
\end{itemize}

\hypertarget{references}{%
\subsubsection{References}\label{references}}

\hypertarget{data-sheets}{%
\paragraph{Data sheets}\label{data-sheets}}

\begin{itemize}
\tightlist
\item
  \href{https://www.invensense.com/wp-content/uploads/2015/02/PS-MPU-9250A-01-v1.1.pdf}{MPU-9250}
  and
  \href{https://cdn.sparkfun.com/assets/learn_tutorials/5/5/0/MPU-9250-Register-Map.pdf}{its
  register map}.
\item
  \href{https://www.akm.com/akm/en/file/datasheet/AK8963C.pdf}{3-axis
  Electronic Compass on IMU}.
\item
  \href{http://www.jnhuamao.cn/bluetooth40_en.zip}{HM-10 (Bluetooth
  breakout module \& firmware)}. The ZIP contains the original data
  sheet, as well as other documentation.
  \href{http://fab.cba.mit.edu/classes/863.15/doc/tutorials/programming/bluetooth/bluetooth40_en.pdf}{MIT
  has a PDF} available, which may me out of date, but is easier to get
  to.
\item
  \href{http://ww1.microchip.com/downloads/en/DeviceDoc/60001168J.pdf}{PIC32MX250}.
\end{itemize}

\hypertarget{code-and-designs-borrowed-from-others}{%
\paragraph{Code and designs borrowed from
others}\label{code-and-designs-borrowed-from-others}}

\begin{itemize}
\tightlist
\item
  \href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/index.html}{Self-Balancing
  Robot} by Desmond Caulley (dc686@cornell.edu), Nadav Nehoran
  (nn233@cornell.edu), Sherry Zhao (hz263@cornell.edu). In particular,
  we borrowed extensively from their
  \href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/dc686_nn233_hz263/i2c_helper.h}{i2c\_helper.h}.
\end{itemize}

\hypertarget{other}{%
\paragraph{Other}\label{other}}

\begin{itemize}
\tightlist
\item
  \href{http://people.ece.cornell.edu/land/courses/ece4760/}{ECE 4760
  Website} by Bruce Land
\item
  Website template:
  \url{https://github.com/tajmone/pandoc-goodies/tree/master/templates/html5/github}.
\end{itemize}

\end{document}

### 2018-02-01 article/article.tex
delete unnecessary stuff

@@
 
-\hypertarget{video}{%
-\subsubsection{Video}\label{video}}
-
 \hypertarget{design}{%
@@
 
-\hypertarget{ieee-code-of-ethics}{%
-\subsubsection{IEEE Code of Ethics}\label{ieee-code-of-ethics}}
-
-Our project, as well as the design and creation project thereof, adheres
-to the \href{https://www.ieee.org/about/ethics.html}{IEEE Code of
-Ethics}. The project itself is relatively small and physically
-self-contained; it did not have any immediate impact on the environment
-or safety and health of potential users, we the creators, or any others.
-We chose the project's topic out of interest and for a challenge,
-without any conflict of interest. Our report is honest and realistic;
-none of our data is fabricated, and our conclusions reflect our hard
-findings. We did not accept any form of bribery.
-
-The project investigated the use of an emerging and popular technology
-(BLE) in a new application (distance sensing). We consistently received
-and built upon advice from our TAs and professor Bruce Land, and cite
-the work we used in the basis of our project.
-
-Additionally, neither we nor our project presented any act of
-discrimination, and our project does not injure others in any way. We
-assisted our colleagues (both fellow group members and other classmates)
-when the occasion arose, such as by offering help debugging or lending a
-spare module before a demo.
-
-Overall, our project complies with the IEEE Policies, 7.8 Code of
-Ethics.
-
 \hypertarget{appendices}{%
@@
 
-\hypertarget{appendix-a-legal-considerations}{%
-\subsubsection{Appendix A: Legal
-Considerations}\label{appendix-a-legal-considerations}}
-
-\begin{itemize}
-\tightlist
-\item
-  The group approves this report for inclusion on the course website.
-\item
-  The group approves the video for inclusion on the course youtube
-  channel.
-\item
-  The TI CC2541 meets FCC CFR47 Part 15 requirements. \footnote{See the
-    top of page 8 of \url{http://www.ti.com/lit/ds/symlink/cc2541.pdf}
-    for compliance information.}
-\end{itemize}
-
-\hypertarget{appendix-b-source-listing}{%
-\subsubsection{Appendix B: Source
-Listing}\label{appendix-b-source-listing}}
-
-Our source code can be found
-\href{https://github.com/orangeturtle739/bluehunters}{on Github}. Each
-file can also be found below:
-
-\hypertarget{code}{%
-\paragraph{Code}\label{code}}
-
-\begin{itemize}
-\tightlist
-\item
-  \href{generated/ble.c.html}{\texttt{ble.c}}
-\item
-  \href{generated/ble.h.html}{\texttt{ble.h}}
-\item
-  \href{generated/imu.c.html}{\texttt{imu.c}}
-\item
-  \href{generated/imu.h.html}{\texttt{imu.h}}
-\item
-  \href{generated/servo.c.html}{\texttt{servo.c}}
-\item
-  \href{generated/servo.h.html}{\texttt{servo.h}}
-\item
-  \href{generated/pt_cornell_1_2_2.c.html}{\texttt{pt\_cornell\_1\_2\_2.c}}
-\item
-  \href{generated/pt_cornell_1_2_2.h.html}{\texttt{pt\_cornell\_1\_2\_2.h}}
-\item
-  \href{generated/main.c.html}{\texttt{main.c}}
-\item
-  \href{generated/main.h.html}{\texttt{main.h}}
-\item
-  \href{generated/config_1_2_2.h.html}{\texttt{config\_1\_2\_2.h}}
-\end{itemize}
-
-\hypertarget{cad}{%
-\paragraph{CAD}\label{cad}}
-
-\hypertarget{sources}{%
-\subparagraph{Sources}\label{sources}}
-
-\begin{itemize}
-\tightlist
-\item
-  \href{generated/main.scad.html}{\texttt{main.scad}}
-\item
-  \href{generated/frame.scad.html}{\texttt{frame.scad}}
-\item
-  \href{generated/wheel.scad.html}{\texttt{wheel.scad}}
-\item
-  \href{generated/drag.scad.html}{\texttt{drag.scad}}
-\item
-  \href{generated/concept.scad.html}{\texttt{concept.scad}}
-\end{itemize}
-
-\hypertarget{stl}{%
-\subparagraph{STL}\label{stl}}
-
-\begin{itemize}
-\tightlist
-\item
-  \href{stl/frame.stl}{\texttt{frame.stl}}
-\item
-  \href{stl/wheel.stl}{\texttt{wheel.stl}}
-\item
-  \href{stl/drag.stl}{\texttt{drag.stl}}
-\end{itemize}
+\hypertarget{appendix-c-schematics}{%
+\subsubsection{Appendix A: Data}\label{appendix-c-schematics}}
 
@@
 \hypertarget{appendix-c-schematics}{%
-\subsubsection{Appendix C: Schematics}\label{appendix-c-schematics}}
+\subsubsection{Appendix B: Schematics}\label{appendix-c-schematics}}
 
@@
 
-\hypertarget{appendix-e-work-distribution}{%
-\subsubsection{Appendix E: Work
-Distribution}\label{appendix-e-work-distribution}}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Jacob:}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Bluetooth firmware hacking
-  \item
-    Bluetooth UART code
-  \item
-    CAD design \& 3D printing
-  \end{itemize}
-\item
-  \textbf{Jane:}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    I2C IMU code
-  \item
-    Robot motion code and gradient descent algorithm
-  \end{itemize}
-\item
-  \textbf{Justin:}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Servo control code
-  \item
-    Board assembly (3 small boards and our 3 of our protoboards)
-  \item
-    Schematic
-  \end{itemize}
-\end{itemize}
-
 \hypertarget{references}{%

### 2018-02-01 article/article.tex
Fluff introduction + remove local refs

@@
 
-We built 2 small cars which used Bluetooth Received Signal Strength
+Position estimation and navigation for robots is an interesting problem
+in the field of robotics. We investigate the challenge of determining
+relative position using signal strength; by placing a moving robot on
+the same plane as a beacon, oriented and positioned randomly, the
+robot can hunt out and move towards the beacon by itself. We discuss 
+our procedures and challenges we encountered in this article.
+
+Specifically, we built 2 small cars which used Bluetooth Received Signal Strength
 Indicator (RSSI) measurements to navigate towards a stationary base
@@
 \subsection{Design}\label{design}}
+
+The design of the project can be split into 3 main parts: the chassis, the electronics (hardware), and the software. The following sections discuss implementation details.
 

### 2018-02-01 article/article.tex
start bib

@@
 \usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
 \usepackage[T1]{fontenc}
@@
 \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
 % use upquote if available, for straight quotes in verbatim environments
@@
 The PIC commmunicates with the IMU via I2C. The IMU includes a breakout
-board for the QFN MPU-9250 module, which itself includes 2 dies. One
+board for the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module, which itself includes 2 dies. One
 contains the 3-axis gyroscope and 3-axis accelerometer, which were not
@@
 \item
-  \href{https://www.invensense.com/wp-content/uploads/2015/02/PS-MPU-9250A-01-v1.1.pdf}{MPU-9250}
-  and
-  \href{https://cdn.sparkfun.com/assets/learn_tutorials/5/5/0/MPU-9250-Register-Map.pdf}{its
-  register map}.
-\item
   \href{https://www.akm.com/akm/en/file/datasheet/AK8963C.pdf}{3-axis
@@
 
+\bibliography{references}{}
+\bibliographystyle{ieeetr}
+
 \end{document}

### 2018-02-02 article/article.tex
Local refs budget
clean up all references

@@
 chips. We realized they were fakes when they did not behave according to
-the Jnhuamao data sheet (see \protect\hyperlink{data-sheets}{the data
-sheets section}). Luckily, the hardware on the fake chips is the same as
+the Jnhuamao data sheet. \footnote{The original data sheet comes from \cite{jnhuamaodatasheet}. A PDF is available at \cite{jnhuamaomit}, which may be out of date, but is eaiser to get.}
+Luckily, the hardware on the fake chips is the same as
 that of the genuine chips, minus an external crystal, and the genuine
 firmware checks for the presence of the crystal, and works even without
-it. \footnote{\url{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}.}
+it. \cite{crystal}
 As such, we reprogrammed the chips with the genuine firmware according
@@
 \href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
-forum post}:
+forum post}  \cite{crystal}:
 
@@
 board was version 540, but Jnhuamao had (at the time we did this
-project)
-\href{http://www.jnhuamao.cn/rom/HMSoft-10-2541-V603.zip}{released
-version 603}. \footnote{Jnhuamao's firmware download page
-  \url{http://www.jnhuamao.cn/download_rom_en.asp?id=1\#}.} They also
-provide
-\href{http://www.jnhuamao.cn/HowToUpgradeFirmware_en.zip}{instructions
-on how to upgrade the firmware}. The basic process is:
+project) released
+version 603. \cite{jnhuamao603} Based on their instructions \cite{jnhuamaoinstructions}, we flashed our chips:
 
@@
 \begin{verbatim}
-OK+DISC:00000000:00000000000000000000000000000000:0000000000:6832A3801EBE:-080
+OK+DISC:00000000:00000000000000000000000000000000:
+  0000000000:6832A3801EBE:-080
 \end{verbatim}
@@
 used in this project, and the other die is the AK8963 3-axis
-magnetometer (compass).
+magnetometer (compass). \cite{ak8963cdatasheet}
 
@@
 configuring the IMU.
+
+We based our IMU code off of the \texttt{i2c\_helper.h} from the Self-Balancing Robot project. \cite{selfbalancingrobot}
 
@@
 where \(A\) and \(n\) are RF propagation parameters in dBm, \(d\) is
-distance in meters, and RSSI is the measured RSSI in dBm. \footnote{L.
-  Peneda, A. Azenha and A. Carvalho, ``Trilateration for indoors
-  positioning within the framework of wireless communications,'' 2009
-  35th Annual Conference of IEEE Industrial Electronics, Porto, 2009,
-  pp.~2732-2737.
-
-  doi: 10.1109/IECON.2009.5415423
-
-  URL:
-  \url{http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=5415423\&isnumber=5414636}}
+distance in meters, and RSSI is the measured RSSI in dBm. \cite{5415423}
 We experimented with RSSI measurements to determine how well they worked
@@
 Both of these are reasonable given the operating range of BLE 4.0.
-
-We reused some of the code from the
-\href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/index.html}{Self-Balancing
-Robot} as
-\protect\hyperlink{code-and-designs-borrowed-from-others}{cited} in our
-references section, and explained in \protect\hyperlink{imu}{IMU}. The
-rest of the design is ours, and is not an infringement on intellectual
-property. We also used the Bluetooth firmware flashing technique we
-discovered on the
-\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
-forums}.
-
-We did not reverse-engineer any technology, and did not encounter any
-issues with trademarks or patents. We did not have to sign
-non-disclosure to get a sample part. We hope to further polish our
-findings and report, and will aim to publish them in a print magazine,
-ideally as a part of ECE 4920.
 
@@
 
-\hypertarget{references}{%
-\subsubsection{References}\label{references}}
-
-\hypertarget{data-sheets}{%
-\paragraph{Data sheets}\label{data-sheets}}
-
-\begin{itemize}
-\tightlist
-\item
-  \href{https://www.akm.com/akm/en/file/datasheet/AK8963C.pdf}{3-axis
-  Electronic Compass on IMU}.
-\item
-  \href{http://www.jnhuamao.cn/bluetooth40_en.zip}{HM-10 (Bluetooth
-  breakout module \& firmware)}. The ZIP contains the original data
-  sheet, as well as other documentation.
-  \href{http://fab.cba.mit.edu/classes/863.15/doc/tutorials/programming/bluetooth/bluetooth40_en.pdf}{MIT
-  has a PDF} available, which may me out of date, but is easier to get
-  to.
-\item
-  \href{http://ww1.microchip.com/downloads/en/DeviceDoc/60001168J.pdf}{PIC32MX250}.
-\end{itemize}
-
-\hypertarget{code-and-designs-borrowed-from-others}{%
-\paragraph{Code and designs borrowed from
-others}\label{code-and-designs-borrowed-from-others}}
-
-\begin{itemize}
-\tightlist
-\item
-  \href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/index.html}{Self-Balancing
-  Robot} by Desmond Caulley (dc686@cornell.edu), Nadav Nehoran
-  (nn233@cornell.edu), Sherry Zhao (hz263@cornell.edu). In particular,
-  we borrowed extensively from their
-  \href{https://people.ece.cornell.edu/land/courses/ece4760/FinalProjects/f2015/dc686_nn233_hz263/final_project_webpage_v2/dc686_nn233_hz263/dc686_nn233_hz263/i2c_helper.h}{i2c\_helper.h}.
-\end{itemize}
-
-\hypertarget{other}{%
-\paragraph{Other}\label{other}}
-
-\begin{itemize}
-\tightlist
-\item
-  \href{http://people.ece.cornell.edu/land/courses/ece4760/}{ECE 4760
-  Website} by Bruce Land
-\item
-  Website template:
-  \url{https://github.com/tajmone/pandoc-goodies/tree/master/templates/html5/github}.
-\end{itemize}
-
 \bibliography{references}{}

### 2018-02-02 article/article.tex
start rewriting

@@
 \hypersetup{
-            pdftitle={Blue Hunters: Bluetooth RSSI Locator Robots},
-            pdfauthor={Jacob Glueck (jng55); Jane Du (zd53); Justin Cray (jgc232)},
-            pdfborder={0 0 0},
-            breaklinks=true}
+    colorlinks=true,
+    linkcolor=blue,
+    filecolor=magenta,
+    urlcolor=cyan,
+}
 \urlstyle{same}  % don't use monospace font for urls
@@
 \graphicspath{{../assets/}{../renderings/}}
+\usepackage{subfig}
+\usepackage{float}
 \makeatletter
@@
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
 \title{Blue Hunters: Bluetooth RSSI Locator Robots}
-\providecommand{\subtitle}[1]{}
-\subtitle{\emph{ECE 4760, Fall 2017}}
 \author{Jacob Glueck (\href{mailto:jng55@cornell.edu}{jng55}) \and Jane Du (\href{mailto:zd53@cornell.edu}{zd53}) \and Justin Cray (\href{mailto:jgc232@cornell.edu}{jgc232})}
@@
 
-Position estimation and navigation for robots is an interesting problem
-in the field of robotics. We investigate the challenge of determining
-relative position using signal strength; by placing a moving robot on
-the same plane as a beacon, oriented and positioned randomly, the
-robot can hunt out and move towards the beacon by itself. We discuss
-our procedures and challenges we encountered in this article.
+Position estimation and navigation for indoor robots is an interesting problem.
+In this article, we investigate using the Received Signal Strength Indicator (RSSI) of Bluetooth Low Energy (BLE) 4.0 chips to allow wheeled mobile robots to navigate towards a stationary base station.
+Each robot was controlled by a PIC32MX250 microcontroller, and used a 3 axis magnetometer as as compass in order to reliably turn, as well as 2 micro 9 g servos to drive.
+Each unit was powered with 3 AA batteries.
+Finally, the chassis and wheels of each car were 3D printed.
+Figure \ref{fig:robotsystem} shows the entire system.
 
-Specifically, we built 2 small cars which used Bluetooth Received Signal Strength
-Indicator (RSSI) measurements to navigate towards a stationary base
-station. The cars and base station used a Bluetooth Low Energy (BLE) 4.0
-module to take the measurements and a PIC32MX250 microcontroller. The
-cars also used a 3 axis magnetometer as as compass in order to reliably
-turn, as well as 2 micro 9 g servos to drive. Each unit was powered with
-3 AA batteries. Finally, the chassis and wheels of each car were 3D
-printed.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{full_system.jpg}
+  \caption{Full system with 2 robots and the base station}
+  \label{fig:robotsystem}
+\end{figure}
 
@@
 \subsection{Design}\label{design}}
-
-The design of the project can be split into 3 main parts: the chassis, the electronics (hardware), and the software. The following sections discuss implementation details.
 
@@
 The robots are made from 4 3D printed pieces: 2 wheels, the frame, and
-the caster in the back. The servos, a 3 AA battery holder, and a
-perfboard containing all the circuitry are mounted directly to the
-frame.
-
-The robots where designed in \href{http://www.openscad.org/}{OpenSCAD},
-and their source code is available in
-\href{https://github.com/orangeturtle739/bluehunters/tree/master/cad}{our
-git repository} and
-\protect\hyperlink{appendix-b-source-listing}{below}. There are three
-files, \href{generated/frame.scad.html}{\texttt{frame.scad}},
-\href{generated/drag.scad.html}{\texttt{drag.scad}}, and
-\href{generated/wheel.scad.html}{\texttt{wheel.scad}}, for each of the
-three parts. The main modules are defined in
-\href{generated/main.scad.html}{\texttt{main.scad}}, and the other files
-just instantiate them. The following renderings show each part:
+the caster in the back.
+Figure \ref{fig:robotchassis} shows the main components.
+The servos, a 3 AA battery holder, and a perfboard containing all the circuitry are mounted directly to the frame.
 
@@
   \centering
+  \subfloat[Frame] {
     \includegraphics[width=0.5\textwidth,height=\textheight]{frame.png}
-\caption{Robot chassis}
+    \label{fig:robotfrane}
+  }
+  \subfloat[Wheel]{
+    \includegraphics[width=0.5\textwidth,height=\textheight]{wheel.png}
+    \label{fig:robotwheel}
+  }
+  \\
+  \subfloat[Caster]{
+    \includegraphics[width=0.5\textwidth,height=\textheight]{drag.png}
+    \label{fig:robotcaster}
+  }
+  \caption{Components of robot chassis}
+  \label{fig:robotchassis}
 \end{figure}
+
+The parts were printed in ABS using Maker Select 3D Printer v2 printers.
+All parts were printed with a layer height of 0.3 mm, as there was no need for a smooth finish or high tolerances.
+The parts where sliced with Cura.
 
@@
   \centering
-\includegraphics[width=0.5\textwidth,height=\textheight]{drag.png}
-\caption{Caster wheel on back of robot}
+  \subfloat[Top view]{
+    \includegraphics[width=0.5\textwidth]{top_view.jpg}
+    \label{fig:robottop}
+  }
+  \subfloat[Side view]{
+    \includegraphics[width=0.5\textwidth]{side_view.jpg}
+    \label{fig:robotside}
+  }
+  \\
+  \subfloat[Front view]{
+    \includegraphics[width=0.5\textwidth]{front_view.jpg}
+    \label{fig:robotfront}
+  }
+  \subfloat[Back view]{
+    \includegraphics[width=0.5\textwidth]{inside_view.jpg}
+    \label{fig:robotback}
+  }
+  \caption{Multiple views of the completed robots}
+  \label{fig:robotviews}
 \end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth,height=\textheight]{wheel.png}
-\caption{Robot wheel}
-\end{figure}
-
-The parts were printed in ABS using
-\href{https://www.monoprice.com/product?c_id=107\&cp_id=10724\&cs_id=1072403\&p_id=13860\&seq=1\&format=2}{Maker
-Select 3D Printer v2} printers. All parts were printed with a layer
-height of 0.3 mm, as there was no need for a smooth finish or high
-tolerances. The parts where sliced with
-\href{https://ultimaker.com/en/products/ultimaker-cura-software}{Cura}.
-
-\hypertarget{pictures}{%
-\subsubsection{Pictures}\label{pictures}}
-
-Full system (base station and 2 robots):
-
-\includegraphics[width=0.5\textwidth,height=\textheight]{full_system.jpg}
-
-Multiple views of the robots:
-
-\begin{longtable}[]{@{}ll@{}}
-\toprule
-\includegraphics{top_view.jpg} &
-\includegraphics{side_view.jpg}\tabularnewline
-\includegraphics{front_view.jpg} &
-\includegraphics{inside_view.jpg}\tabularnewline
-\bottomrule
-\end{longtable}
 
@@
 
-\hypertarget{bluetooth-modules}{%
-\paragraph{Bluetooth modules}\label{bluetooth-modules}}
+Each robot consisted of 4 main components:
+\begin{enumerate}
+  \item PIC32MC250F128B microcontroller.
+  \item MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer.
+  \item HM-10 BLE module with a Texas Instruments CC2541 chip.
+  \item Two 9 gram FS90R micro continuous rotation servos.
+\end{enumerate}
 
-The HM-10 bluetooth modules we bought off Ebay were fakes: they were not
-made by Jnhuamao \footnote{Jnhuamao is the company which makes the HM-10
-  module, along with other Bluetooth modules. Their English website is:
-  \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}, and did not come
-with genuine Jnhuamao firmware. However, they did have real TI CC2541
-chips. We realized they were fakes when they did not behave according to
-the Jnhuamao data sheet. \footnote{The original data sheet comes from \cite{jnhuamaodatasheet}. A PDF is available at \cite{jnhuamaomit}, which may be out of date, but is eaiser to get.}
-Luckily, the hardware on the fake chips is the same as
-that of the genuine chips, minus an external crystal, and the genuine
-firmware checks for the presence of the crystal, and works even without
+The components are connected as shown in figure \ref{fig:schematic}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{schematic.png}
+  \caption{Schematic}
+  \label{fig:schematic}
+\end{figure}
+
+\paragraph{Bluetooth modules}
+
+The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao,
+\footnote{Jnhuamao is the company which makes the HM-10 module, along with other Bluetooth modules. Their English website is: \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}
+and did not come with genuine Jnhuamao firmware.
+However, they did have genuine TI CC2541 chips.
+We realized they were fakes when they did not behave according to
+the Jnhuamao data sheet.
+\footnote{The original data sheet comes from \cite{jnhuamaodatasheet}.
+A PDF is available at \cite{jnhuamaomit}, which may be out of date, but is eaiser to get.}
+Luckily, the hardware on the fake chips is the same as that of the genuine chips, minus an external crystal, and the genuine firmware checks for the presence of the crystal, and works even without
 it. \cite{crystal}
 As such, we reprogrammed the chips with the genuine firmware according
-to an
-\href{http://forum.arduino.cc/index.php?topic=393655.msg2709528\#msg2709528}{Arduino
-forum post}  \cite{crystal}:
+to an Arduino \cite{crystal}:
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
 \item
-  We soldered wires to the programming pins on the breakout boards, and
-  connected those pins to an
-  \href{https://www.pjrc.com/teensy/teensy31.html}{Arduino Teensy 3.2}.
-  We chose a Teensy because it is 3.3 V as opposed to 5, which would
-  damage the 3.3 V CC2541.
-
-  The pins were connected as follows:
+  We soldered wires to the programming pins on the breakout boards, and connected those pins to an Arduino Teensy 3.2.
+  We chose a Teensy because it is 3.3 volts as opposed to 5, which would damage the 3.3 volt CC2541. The pins were connected as follows:
 
@@
 
-  The layout of the HM-10 board is:
+  Figure \ref{fig:hm10} shows the layout of the HM-10 module.
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.3\textwidth]{hm10_pins.png}
+    \caption{HM-10 module pin layout \cite{crystal}}
+    \label{fig:hm10}
+  \end{figure}
 
-  \includegraphics{hm10_pins.png}
 \item
-  We uploaded the
-  \href{https://github.com/RedBearLab/CCLoader/blob/master/Arduino/CCLoader/CCLoader.ino}{\texttt{CCLoader.ino}
-  sketch} to the Arduino.
+  We uploaded the \texttt{CCLoader.ino} \cite{ccloader} sketch to the Arduino.
 \item
-  Finally, we ran (in a Windows virtual machine)
-  \href{https://github.com/RedBearLab/CCLoader/tree/master/Windows}{\texttt{CCLoader.exe}}.
+  Finally, we ran (in a Windows virtual machine, due to the dubious origin of the software) \texttt{CCLoader.exe} \cite{ccloaderexe}.
   This program takes 3 arguments:
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\ExtensionTok{CCLoader.exe} \OperatorTok{<}\NormalTok{COM Port}\OperatorTok{>} \OperatorTok{<}\NormalTok{Firmware.bin}\OperatorTok{>}\NormalTok{ 0}
-\end{Highlighting}
-\end{Shaded}
+  \begin{Verbatim}[gobble=4]
+    CCLoader.exe <COM Port> <Firmware.bin> 0
+  \end{Verbatim}
 
-  The firmware file came from the same Arduino form post, and can be
-  found
-  \href{http://forum.arduino.cc/index.php?action=dlattach;topic=393655.0;attach=183702}{here}.
+  The firmware file came from the Arduino forum post \cite{crystal}, and can be
+  found at \url{http://forum.arduino.cc/index.php?action=dlattach;topic=393655.0;attach=183702}.
 \end{enumerate}
 
-There is also an excellent
-\href{https://www.youtube.com/watch?v=ez3491-v8Og}{YouTube video}, which
-explains the firmware flashing process.
-
-After flashing genuine firmware onto the chip, the next step was to
-update the firmware to the latest version. The firmware flashed onto the
-board was version 540, but Jnhuamao had (at the time we did this
-project) released
-version 603. \cite{jnhuamao603} Based on their instructions \cite{jnhuamaoinstructions}, we flashed our chips:
+After flashing genuine firmware onto the chip, the next step was to update the firmware to the latest version.
+The firmware flashed onto the board was version 540, but Jnhuamao had (at the time we did this project) released version 603. \cite{jnhuamao603}
+Based on their instructions \cite{jnhuamaoinstructions}, we flashed our chips:
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
@@
 
-\hypertarget{software}{%
-\subsubsection{Software}\label{software}}
+\subsubsection{Software}
 
@@
 
-\hypertarget{ble}{%
-\paragraph{BLE}\label{ble}}
+\paragraph{BLE}
 
@@
 
-\hypertarget{imu}{%
-\paragraph{IMU}\label{imu}}
+\paragraph{IMU}
 
@@
 
-\hypertarget{servos}{%
-\paragraph{Servos}\label{servos}}
+\paragraph{Servos}
 
@@
 
-\hypertarget{protothreads}{%
-\paragraph{Protothreads}\label{protothreads}}
+\paragraph{Protothreads}
 
@@
 
-\hypertarget{gradient-descent}{%
-\paragraph{Gradient descent}\label{gradient-descent}}
+\paragraph{Gradient descent}
 
@@
 
-\hypertarget{testing}{%
-\subsubsection{Testing}\label{testing}}
+\subsubsection{Testing}
 
-\hypertarget{ble-1}{%
-\paragraph{BLE}\label{ble-1}}
+\paragraph{BLE}
 
@@
 
-\hypertarget{imu-1}{%
-\paragraph{IMU}\label{imu-1}}
+\paragraph{IMU}
 
@@
 
-\hypertarget{servos-1}{%
-\paragraph{Servos}\label{servos-1}}
+\paragraph{Servos}
 
@@
 
-\hypertarget{whole-system}{%
-\paragraph{Whole system}\label{whole-system}}
+\paragraph{Whole system}
 
@@
 
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
+\subsection{Results}
 
@@
 
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
+\subsection{Conclusions}
 
@@
 
-\hypertarget{appendices}{%
-\subsection{Appendices}\label{appendices}}
-
-\hypertarget{appendix-c-schematics}{%
-\subsubsection{Appendix A: Data}\label{appendix-c-schematics}}
+\subsubsection{Appendix A: Data}
 
@@
 
-\hypertarget{appendix-c-schematics}{%
-\subsubsection{Appendix B: Schematics}\label{appendix-c-schematics}}
+\subsubsection{Appendix B: Bill Of Materials}
 
-\includegraphics{schematic.png}
-
-\hypertarget{appendix-d-bill-of-materials}{%
-\subsubsection{Appendix D: Bill Of
-Materials}\label{appendix-d-bill-of-materials}}
+This project was done with a \$150 budget in mind.
+Costs for parts labeled ``Lab rental'' were borrowed from the department of Electrical and Computer Engineering at Cornell University, but can also be found at vendors like Digikey, Adafruit, and Amazon.
 

### 2018-02-02 article/article.tex
finish first round of edits

@@
 
-The robots were programmed in C, using the
-\href{http://www.microchip.com/mplab/mplab-x-ide}{MPLAB X IDE} v4.0,
-with the \href{http://www.microchip.com/mplab/compilers}{XC32} v1.4
-compiler and the
-\href{http://www.microchip.com/SWLibraryWeb/product.aspx?product=PIC32\%20Peripheral\%20Library}{PIC
-32 Legacy Peripheral Library (plib)}. The full source code is available
-in
-\href{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}{our
-git repository} and
-\protect\hyperlink{appendix-b-source-listing}{below}. The code is
-divided into 5 main units:
-
-\begin{itemize}
-\tightlist
-\item
-  \href{generated/ble.c.html}{\texttt{ble.c}}: contains all the
-  functions for interacting with the BLE device over UART.
-\item
-  \href{generated/imu.c.html}{\texttt{imu.c}}: contains all the
-  functions for interacting with the IMU (which contained the
-  magnetometer) over I2C.
-\item
-  \href{generated/servo.c.html}{\texttt{servo.c}}: contains all the
-  functions for interacting with the servos using PWM.
-\item
-  \href{generated/main.c.html}{\texttt{main.c}}: the main code.
-\item
-  \href{generated/pt_cornell_1_2_2.c.html}{\texttt{pt\_cornell\_1\_2\_2.c}}:
-  contains the functions which were originally declared in the
-  protothreads header file. However, since they were in the header file,
-  if multiple source files included the header file, there would be
-  linking errors due to duplicate definitions of symbols. Moving the
-  protothreads definitions to a separate file resolved this issue.
-\end{itemize}
+The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
+compiler and the PIC 32 Legacy Peripheral Library (plib).
+The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}.
 
@@
 the PIC for communicating with the BLE module, and used \texttt{UART2}
-for communicating with the computer (for debugging). The BLE files
-define a series of macros and \texttt{PT\_THREAD} functions for printing
-characters and reading lines in a non-blocking fashion using
-protothreads. All of the \texttt{PT\_THREAD} functions can be spawned
-with \texttt{PT\_SPAWN} to run them.
-
-The \href{generated/main.c.html}{\texttt{main.c}} file contains a thread
-which does all the communication with the Bluetooth chip using these
-functions. The commands it sends to the chip are:
+for communicating with the computer (for debugging).
+We used the following commands to interact with the bluetooth chip:
 
@@
 
-The PIC commmunicates with the IMU via I2C. The IMU includes a breakout
-board for the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module, which itself includes 2 dies. One
-contains the 3-axis gyroscope and 3-axis accelerometer, which were not
+The PIC communicates with the IMU via I2C.
+The IMU includes a breakout board for the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module, which itself includes 2 dies.
+One contains the 3-axis gyroscope and 3-axis accelerometer, which were not
 used in this project, and the other die is the AK8963 3-axis
@@
 
-It is connected to the rest of the MPU module via an auxillary I2C bus,
+It is connected to the rest of the MPU module via an auxiliary I2C bus,
 so it is not connected to the MPU's main I2C bus by default. While the
@@
 IMU, the compass also needs pass-through mode to be enabled on the IMU
-to make it an accessible slave on the I2C bus.
-
-Some useful functions as defined in
-\href{generated/imu.h.html}{\texttt{imu.h}} are described below:
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{void\ imu\_init()}: Initializes the MPU-9250, including
-  configuring the chip to allow reading the compass.
-\item
-  \texttt{int\ imu\_get\_heading()} Returns the heading of the robot as
-  a value between -180 and 180. The compasses were not completely
-  calibrated to find magnetic north; it only ensures angles are correct
-  relevant to past headings.
-\item
-  \texttt{void\ imu\_mag\_read\_data(int\ *\ destination)} Fetches
-  compass readings; saves register values into \texttt{destination} in
-  the form \texttt{{[}x,\ y,\ z{]}}.
-\item
-  \texttt{int\ angle\_diff(int\ source,\ int\ target)} Gets the
-  difference between two angles in degrees to account for discontinuity
-  between -180 and 180 degrees.
-\item
-  \texttt{int\ degree(int\ deg)}: Offsets degree values so that they
-  fall within the range -180 to 180 degrees.
-\end{itemize}
-
-Initializing the IMU involves opening the I2C module, and then
-configuring the IMU.
-
-We based our IMU code off of the \texttt{i2c\_helper.h} from the Self-Balancing Robot project. \cite{selfbalancingrobot}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  We open the \texttt{I2C2} module; \texttt{I2C1} uses pins already used
-  by the connections to the Bluetooth module. We open it with the baud
-  rate generator value
-  \texttt{BRG\ =\ (Fpb\ /\ 2\ /\ baudrate)\ -\ 2\ =\ 4e7\ /\ 2\ /\ 4e5\ -\ 2\ =\ 48},
-  as specified for \texttt{OpenI2C2()} in the
-  \protect\hyperlink{references}{peripheral libraries}.
-\item
-  Pass through is enabled, interrupts for data ready are enabled, the
-  IMU as an I2C master function is disabled, and the sensor is powered
-  up.
-\end{enumerate}
-
-This enables the PIC to talk to the AK8963. The AK8963 has several modes
+to make it an accessible slave on the I2C bus. We based our IMU code off of the \texttt{i2c\_helper.h} from the Self-Balancing Robot project. \cite{selfbalancingrobot} The AK8963 has several modes
 of operation, and the chip must be set to power-down mode before
 switching to other modes. We read compass values with single measurement
-mode, as specified below:
+mode, as in figure \ref{fig:imu_single_measurement}.
 
-\includegraphics{imu_single_measurement.png}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{imu_single_measurement.png}
+  \caption{IMU single measurement mode}
+  \label{fig:imu_single_measurement}
+\end{figure}
 
+In order to read the magnetometer, we:
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
 \item
@@
 
-Additional helper methods used in I2C were defined in
-\href{generated/imu.c.html}{\texttt{imu.c}}. These are largely based on
-the files from the
-\protect\hyperlink{code-and-designs-borrowed-from-others}{self-balancing
-robot}.
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{char\ i2c\_read\_device(char\ device,\ char\ address)} Reads
-  the data from a single register at \texttt{address}
-\item
-  \texttt{void\ i2c\_write\_byte(char\ device,\ char\ address,\ char\ data)}
-  All configurations used in this project involved writing single bytes
-  of data.
-\item
-  \texttt{void\ i2c\_wait(int\ cnt)} Writes 2 nops; reads require time
-  to return a value, and calling reads consecutively.
-\end{itemize}
-
 In order to calibrate the compass, the robots spun in place when powered
@@
 
-For the drive systems, we used the FS90R servos. These servos are
+For the drive system, we used the FS90R servos. These servos are
 continuous rotation servos with a stall torque of 1.5 kg-cm. Continuous
@@
 
-\paragraph{Protothreads}
-
-While the provided protothreads library was very helpful, it did not
-work correctly when used in multiple compilation units. The reason for
-this was that it defined variables and functions in the header file, so
-if multiple source files included
-\href{generated/pt_cornell_1_2_2.h.html}{\texttt{pt\_cornell\_1\_2\_2.h}},
-linking would fail with duplicate symbol definitions. Furthermore, if
-multiple files included
-\href{generated/config_1_2_2.h.html}{\texttt{config\_1\_2\_2.h}},
-linking would fail with the obscure error:
-
-\begin{verbatim}
-section `.config_BFC00BF4' will not fit in region `config2'
-\end{verbatim}
-
-This happened because if multiple compilation units included
-\texttt{\#pragma\ config} directives (which they did by including a
-\texttt{\#include\ "config\_1\_2\_2.h"} directive), then the linker
-would try to assign too many symbols to the configuration region.
-
-The fix to this was twofold:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  We modified protothreads to have both a header and source file, with
-  definitions only in the source file.
-\item
-  We only included
-  \href{generated/config_1_2_2.h.html}{\texttt{config\_1\_2\_2.h}} in
-  the main C file, and moved all macro definitions to the protothreads
-  header.
-\end{enumerate}
-
-The updated versions of protothreads can be found in
-\protect\hyperlink{appendix-b-source-listing}{Appendix B}.
-
 \paragraph{Gradient descent}
@@
 The algorithm for deciding what path to follow is a basic version of
-gradient descent. The following image represents the decision-making
-state machine, where the starting state is \textbf{Measure rssi twice,
-take average}.
+gradient descent. Figure \ref{fig:graddesc} shows 2 versions of the gradient descent algorithm.
 
-\includegraphics{grad_desc.png}
+\begin{figure}
+  \centering
+  \subfloat[Simple gradient descent]{
+    \includegraphics[width=0.4\textwidth]{grad_desc.png}
+    \label{fig:simplegrad}
+  }
+  \subfloat[Compensating gradient descent]{
+      \includegraphics[width=0.5\textwidth]{turn_correction.png}
+    \label{fig:simplegrad}
+  }
+  \caption{2 gradient descent algorithms. The initial state is \emph{measure RSSI twice, take average}.}
+  \label{fig:graddesc}
+\end{figure}
 
-We also implemented and tested the following improved version that
-allows for correction; a car that has just moved forward and detected a
-weakened signal does not know whether the beacon to its left or right.
-If after turning, the signal is still weaker, it has picked the wrong
-turn. This decision process corrects this (same starting state:
-\textbf{Measure rssi twice, take average}):
-
-\includegraphics{turn_correction.png}
-
-It did not prove much more accurate than randomized gradient descent,
-largely due to noisy readings from IMU and Bluetooth signal strength. As
-such, we used the simplified version.
-
-\subsubsection{Testing}
-
-\paragraph{BLE}
-
-Once we flashed the Bluetooth chips with the proper firmware, we used
-the
-\href{https://itunes.apple.com/us/app/lightblue-explorer/id557428110?mt=8}{LightBlue}
-app to test communications. Using the app, we could discover the
-Bluetooth chips, and send and receive data. By initially using a the
-iPhone app, which was known to work, we could debug the Bluetooth chips
-in isolation. Once that worked, we had the chips communicate with each
-other. Finally, we had the chips take RSSI measurements.
-
-We also experimented with RSSI, as described in
-\protect\hyperlink{results}{results}. This data validated the
-effectiveness of RSSI as an approximate metric for distance measurement,
-though confirmed that it would be highly variable.
-
-\paragraph{IMU}
-
-\begin{itemize}
-\tightlist
-\item
-  Basic I2C: Basic testing was required to ensure I2C communication with
-  the chip was working. This included testing the correctness of the I2C
-  read and write functions (including the order of starts, restarts, and
-  idles in accordance with the I2C protocol and the chip datasheets), as
-  well basic parameters like the address of the MPU-9250 and its
-  registers.
-\item
-  Compass I2C: This was more complex, and mostly involved testing the
-  usage of the modes of both the MPU-9250 (IMU module) and the AK8963
-  (the compass itself). Performance testing was also important here, as
-  the compass would not allow rapid reads in quick succession; we
-  inserted nops in the
-  \texttt{imu\_mag\_read\_data(int\ *\ destination)} for this reason.
-\item
-  Compass calibration: We tested incrementally, first using hard-coded
-  maximum and minimum observed values outputted from the IMU, and then
-  with values taken from a self-calibration in which the robot would
-  spin in a circle, and the PIC would record the largest and smallest
-  values it saw on the X- and Y-axes.
-\end{itemize}
-
-\paragraph{Servos}
-
-We tested the functionality of servo code by testing individual
-functionality on both the hardware and software side. We tested the
-servo code in stages, first implementing control at a fixed speed and
-then implementing directional motion and lastly turning. This allowed us
-to verify our parts of the code and the functionality of the servos.
-
-\paragraph{Whole system}
-
-\begin{itemize}
-\tightlist
-\item
-  Some subsystem testing was required for the servo with the IMUs, for
-  turning/driving straight with feedback. We debugged by reading in IMU
-  values via UART and observing the robot's turns.
-\item
-  We tested extensively with the beacon and two hunters in different
-  environments. Most of our initial testing was in a large, open space
-  with a few metal tables on the periphery. We later tested in the lab
-  in the Digital Lab (Phillips 238), and in the hallway where our demo
-  took place. The confined space of the inside of the lab and hallway,
-  as well as the presence of more reflective bodies (metal doors,
-  tables, people) negatively impacted the performance of our hunters,
-  and it was necessary to recalibrate the hunters to work better in the
-  demo environment.
-\end{itemize}
+The purpose of the compensating algorithm was to allow the robots to recover faster after taking a wrong turn: if it just just moved forward and detected a
+weakened signal, turned, and still detected a weaker signal, then it must have turned the wrong way, so it does a 180.
+However, it did not prove much more accurate than randomized gradient descent, largely due to noisy readings from IMU and Bluetooth signal strength.
+As such, we used the simplified version.
 
@@
 square) at a time. At each point, we took 3 RSSI measurements and
-averaged them. The graph below displays the results:
+averaged them. Figure \ref{fig:rssiplot} shows the results (the original data is in table \ref{table:rssidata}).
 
-\includegraphics{rssi-chart.png}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{rssi-chart.png}
+  \caption{Plot showing bluetooth RSSI as a function of distance}
+  \label{fig:rssiplot}
+\end{figure}
+
+\begin{table}
+  \centering
+  \begin{tabular}[]{@{}llllll@{}}
+  \toprule
+  Distance (feet) & RSSI A & RSSI B & RSSI C & RSSI Average & Expected
+  RSSI \\
+  \midrule
+  1 & -48 & -48 & -48 & -48.00 & -48.00 \\
+  2 & -56 & -58 & -52 & -55.33 & -57.03 \\
+  3 & -66 & -66 & -68 & -66.67 & -62.31 \\
+  4 & -64 & -68 & -65 & -65.67 & -66.06 \\
+  5 & -60 & -70 & -64 & -64.67 & -68.97 \\
+  6 & -80 & -80 & -76 & -78.67 & -71.34 \\
+  7 & -80 & -79 & -65 & -74.67 & -73.35 \\
+  8 & -68 & -66 & -72 & -68.67 & -75.09 \\
+  \bottomrule
+  \end{tabular}
+  \caption{RSSI and distance data}
+  \label{table:rssidata}
+\end{table}
 
@@
 measurements, we could not do so precisely -- it seemed to even depend
-on where our feet where! See \protect\hyperlink{rssi-data}{Appendix B}
-for the source data for the table.
+on where our feet where!
 
@@
 
-This assignment was an interesting exploration into short-range distance
-determination using Bluetooth, a generally unconventional approach. We
-were warned beforehand about the difficulties in using Bluetooth RSSI, a
-measurement particularly susceptible to multipath interference, as a
-distance sensor -- both in the papers and reference materials researched
-before implementation, as well as by course staff in the early stages of
-our project. We worked with the course staff to simplify and stratify
-our project into workable milestones. Our final project is a reduced
-version of our original plan.
-
-Our hunting bots worked reliably when they stayed within roughly 1 m of
-the beacon. After this, they entered the land of shallow gradients: the
-signal strength from the beacon (already noisy) would not change very
-much, and often only due to noise. The hunting bots normally could never
-recover from this. Our bots also worked better in larger, more open
-spaces, preferably without metal structures like tables or doors nearby.
-Both of these are reasonable given the operating range of BLE 4.0.
+This project was an interesting exploration into short-range distance
+determination using Bluetooth, a generally unconventional approach.
+We knew that Bluetooth RSSI would be noisy, mostly due to multipath interference.
+The robots worked reliably when they stayed within roughly 1 meter of the beacon.
+After this, they entered the land of shallow gradients:
+the signal strength from the beacon (already noisy) would not change very
+much, and often only due to noise.
+They normally could never recover from this.
 
 There are many potential avenues for improvements or further
-development. In addition to more circumstantial difficulties encountered
-during the project work period (such as extremely unbalanced servos,
-where one always ran faster than the other), two areas of potential
-improvement are as follows:
+development:
 
@@
 \item
-  \emph{More complete usage of IMU}. A lot of development time was spent
-  simply on getting the PIC to compass communication running, and not
-  much time was spent on calibrating the sensor or processing the data.
-  In further stages of this project, we could continue to work towards
-  dead reckoning of the mobile robots. This, paired with inter-swarm
+  \emph{More complete usage of IMU}.
+Additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
+This, paired with inter-swarm
   communication, would make for a much more sophisticated and likely
@@
 
-\subsubsection{Appendix A: Data}
-
-\hypertarget{rssi-data}{%
-\paragraph{RSSI data}\label{rssi-data}}
-
-\begin{longtable}[]{@{}llllll@{}}
-\toprule
-Distance (feet) & RSSI A & RSSI B & RSSI C & RSSI Average & Expected
-RSSI\tabularnewline
-\midrule
-\endhead
-1 & -48 & -48 & -48 & -48.00 & -48.00\tabularnewline
-2 & -56 & -58 & -52 & -55.33 & -57.03\tabularnewline
-3 & -66 & -66 & -68 & -66.67 & -62.31\tabularnewline
-4 & -64 & -68 & -65 & -65.67 & -66.06\tabularnewline
-5 & -60 & -70 & -64 & -64.67 & -68.97\tabularnewline
-6 & -80 & -80 & -76 & -78.67 & -71.34\tabularnewline
-7 & -80 & -79 & -65 & -74.67 & -73.35\tabularnewline
-8 & -68 & -66 & -72 & -68.67 & -75.09\tabularnewline
-\bottomrule
-\end{longtable}
-
-\subsubsection{Appendix B: Bill Of Materials}
+\subsubsection{Appendix A: Bill Of Materials}
 

### 2018-02-04 article/article.tex
incorperate schematics
pass of edits

@@
 \usepackage{graphicx,grffile}
-\graphicspath{{../assets/}{../renderings/}}
+\graphicspath{{../assets/}{../renderings/}{../schematics/}}
 \usepackage{subfig}
 \usepackage{float}
+\usepackage[page]{appendix}
 \makeatletter
@@
 
-Position estimation and navigation for indoor robots is an interesting problem.
 In this article, we investigate using the Received Signal Strength Indicator (RSSI) of Bluetooth Low Energy (BLE) 4.0 chips to allow wheeled mobile robots to navigate towards a stationary base station.
@@
 
-\hypertarget{design}{%
-\subsection{Design}\label{design}}
+\subsection{Hardware}
 
-\hypertarget{chassis}{%
-\subsubsection{Chassis}\label{chassis}}
+\subsubsection{Chassis}
 
@@
   \subfloat[Frame] {
-    \includegraphics[width=0.5\textwidth,height=\textheight]{frame.png}
+    \includegraphics[width=0.4\textwidth,height=\textheight]{frame.png}
     \label{fig:robotfrane}
@@
   \subfloat[Wheel]{
-    \includegraphics[width=0.5\textwidth,height=\textheight]{wheel.png}
+    \includegraphics[width=0.4\textwidth,height=\textheight]{wheel.png}
     \label{fig:robotwheel}
@@
   \subfloat[Caster]{
-    \includegraphics[width=0.5\textwidth,height=\textheight]{drag.png}
+    \includegraphics[width=0.4\textwidth,height=\textheight]{drag.png}
     \label{fig:robotcaster}
@@
 All parts were printed with a layer height of 0.3 mm, as there was no need for a smooth finish or high tolerances.
-The parts where sliced with Cura.
+The parts were designed in OpenSCAD, and sliced with Cura.
 
@@
   \subfloat[Top view]{
-    \includegraphics[width=0.5\textwidth]{top_view.jpg}
+    \includegraphics[width=0.4\textwidth]{top_view.jpg}
     \label{fig:robottop}
@@
   \subfloat[Side view]{
-    \includegraphics[width=0.5\textwidth]{side_view.jpg}
+    \includegraphics[width=0.4\textwidth]{side_view.jpg}
     \label{fig:robotside}
@@
   \subfloat[Front view]{
-    \includegraphics[width=0.5\textwidth]{front_view.jpg}
+    \includegraphics[width=0.4\textwidth]{front_view.jpg}
     \label{fig:robotfront}
@@
   \subfloat[Back view]{
-    \includegraphics[width=0.5\textwidth]{inside_view.jpg}
+    \includegraphics[width=0.4\textwidth]{inside_view.jpg}
     \label{fig:robotback}
@@
 
-The components are connected as shown in figure \ref{fig:schematic}.
+The PIC32 and supporting circuity was on a small PCB (the ``Small Dev Board'' from ECE 4760 at Cornell University, the class during which we built this) shown in figure \ref{fig:schematic-sdb}.
+The rest of the components were assembled on solderable breadboard as shown in figure \ref{fig:schematic-robot}.
 
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{schematic.png}
+  \subfloat[] {
+    \includegraphics[width=0.9\textwidth]{bluehunters-robot.png}
+    \label{fig:schematic-robot}
+  }
+  \\
+  \subfloat[] {
+    \includegraphics[width=0.9\textwidth]{bluehunters-sdb.png}
+    \label{fig:schematic-sdb}
+  }
   \caption{Schematic}
@@
 As such, we reprogrammed the chips with the genuine firmware according
-to an Arduino \cite{crystal}:
+to an Arduino Fourm post \cite{crystal}:
 
@@
 
-  The firmware file came from the Arduino forum post \cite{crystal}, and can be
-  found at \url{http://forum.arduino.cc/index.php?action=dlattach;topic=393655.0;attach=183702}.
+  The firmware file came from the Arduino forum post. \footnote{The firmware file can be found here:  \url{http://forum.arduino.cc/index.php?action=dlattach;topic=393655.0;attach=183702}}
 \end{enumerate}
@@
   Run the \texttt{HMSoft.exe} program distributed in the firmware update
-  download. We ran it in a Windows virtual machine because we did not
-  trust it.
+  download. (Again in a Windows virtual machine, due to the dubious origin of the software.)
 \item
@@
 
-\subsubsection{Software}
+\subsection{Software}
 
 The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
-compiler and the PIC 32 Legacy Peripheral Library (plib).
-The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}.
+compiler and the PIC 32 Legacy Peripheral Library (plib). \footnote{
+The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}}
 
-\paragraph{BLE}
+\subsubsection{BLE}
 
@@
   \texttt{AT+DISI?}: performed only on the robots, causing a discovery
-  scan. The result of the scan is a bunch of lines of the form:
+  scan. The result of the scan is a series of lines of the form:
 
@@
 
-One interesting thing to note about the chip is that commands do not
-have to end with newlines or carriage returns. However, if sent, the
-chip will ignore them.
+One interesting thing to note about the chip is that commands do not have to end with newlines or carriage returns.
+However, if sent, the chip will ignore them.
 
@@
 
-The PIC communicates with the IMU via I2C.
-The IMU includes a breakout board for the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module, which itself includes 2 dies.
-One contains the 3-axis gyroscope and 3-axis accelerometer, which were not
+The PIC communicates with the IMU via I2C. The QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module has 2 dies:
+one contains the 3-axis gyroscope and 3-axis accelerometer, which were not
 used in this project, and the other die is the AK8963 3-axis
@@
 
-It is connected to the rest of the MPU module via an auxiliary I2C bus,
-so it is not connected to the MPU's main I2C bus by default. While the
-accelerometer and gyroscope registers can be read after powering up the
-IMU, the compass also needs pass-through mode to be enabled on the IMU
-to make it an accessible slave on the I2C bus. We based our IMU code off of the \texttt{i2c\_helper.h} from the Self-Balancing Robot project. \cite{selfbalancingrobot} The AK8963 has several modes
-of operation, and the chip must be set to power-down mode before
-switching to other modes. We read compass values with single measurement
-mode, as in figure \ref{fig:imu_single_measurement}.
+The compass is connected to the rest of the MPU module via an auxiliary I2C bus.
+In order to make the compass an accessible slave on the I2C bus, pass-through mode to be enabled on the IMU.
+\footnote{We based our IMU code off of another ECE 4760 project, the Self-Balancing Robot. \cite{selfbalancingrobot} Their \texttt{i2c\_helper.h} was very helpful.}
+The AK8963 has several modes of operation, and the chip must be set to power-down mode before switching to other modes.
+We read compass values with single measurement mode, as in figure \ref{fig:imu_single_measurement}:
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.9\textwidth]{imu_single_measurement.png}
-  \caption{IMU single measurement mode}
-  \label{fig:imu_single_measurement}
-\end{figure}
-
-In order to read the magnetometer, we:
 \begin{enumerate}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\textwidth]{imu_single_measurement.png}
+  \caption{IMU single measurement mode}
+  \label{fig:imu_single_measurement}
+\end{figure}
+
 In order to calibrate the compass, the robots spun in place when powered
-on. They recorded the maximum and minimum values for each axes, and used
+on.
+They recorded the maximum and minimum values for each axes, and used
 that data to scale and center the magnetometer readings.
@@
 
-For the drive system, we used the FS90R servos. These servos are
-continuous rotation servos with a stall torque of 1.5 kg-cm. Continuous
-rotation servos operate based on PWM duty cycle; since servos operate at
-a pulse width of 1 to 2 ms out of 20 ms, we used a timer with a
-prescaler of 32 and a maximum count of 25600, and set the pwm at numbers
-ranging from 1280 to 2560 (corresponding to the right pulse width).
-
-A 50\% duty cycle means the servos are still. A less than 50\% duty
-cycle means the servo turns clockwise and a greater than 50\% duty cycle
-means the servo turns counter-clockwise. To implement the PWM, we use
-the output capture on the PIC. We set the duty cycle using the built in
-timers. We set a max and min duty cycle value which were experimentally
-derived. The middle point is the average of the two. Tank drive is used
-to control the system where both wheels are driven forward or point
-turns where one wheel is turned in one direction and one is turned in
-the other direction.
+For the drive system, we used FS90R servos.
+These servos are continuous rotation servos with a stall torque of 1.5 kg-cm.
+Like most servos, they are driven with PWM signals at 50 hertz (20 millisecond periods).
+A duty cycle with about 1.5 milliseconds on is roughly neutral.
+With the PIC32s running at 40 MHz, a prescaler of 32 and a maximum count of 25600 produces the required 50 hertz signal ($25600/\left(40\cdot 10^6 \frac{1}{\text{s}}/ 32 \right) = 0.02048 \text{ ms}$).
+The PWM values (in timer counts) thus ranged from ranging from 1280 to 2560.
 
@@
       \includegraphics[width=0.5\textwidth]{turn_correction.png}
-    \label{fig:simplegrad}
+    \label{fig:compensatinggrad}
   }
@@
 
-The purpose of the compensating algorithm was to allow the robots to recover faster after taking a wrong turn: if it just just moved forward and detected a
-weakened signal, turned, and still detected a weaker signal, then it must have turned the wrong way, so it does a 180.
-However, it did not prove much more accurate than randomized gradient descent, largely due to noisy readings from IMU and Bluetooth signal strength.
+The purpose of the compensating algorithm (figure \ref{fig:compensatinggrad}) was to allow the robots to recover faster after taking a wrong turn: if a robot just just moved forward and detected a weaker signal, turned, and still detected a weaker signal, then it must have turned the wrong way, so it does a 180.
+However, it did not prove any more accurate than randomized gradient descent, largely due to noisy RSSI readings.
 As such, we used the simplified version.
@@
 
-We expected that RSSI would vary with distance according to the
-following relation:
-
-\(\text{RSSI} = A - 10 n \log(d)\)
-
-where \(A\) and \(n\) are RF propagation parameters in dBm, \(d\) is
-distance in meters, and RSSI is the measured RSSI in dBm. \cite{5415423}
-We experimented with RSSI measurements to determine how well they worked
-by taking 2 Bluetooth modules, and measuring the RSSI while changing the
-distance between them. One remained stationary on the floor and the
-other was moved away from it 1 floor tile (each floor tile is a 1 foot
-square) at a time. At each point, we took 3 RSSI measurements and
-averaged them. Figure \ref{fig:rssiplot} shows the results (the original data is in table \ref{table:rssidata}).
+We expected that RSSI would vary with distance according to equation \ref{eq:rssi}:
+\begin{equation} \label{eq:rssi}
+  \text{RSSI} = A - 10 n \log(d),
+\end{equation}
+where $A$ and $n$ are RF propagation parameters in dBm, $d$ is distance in meters, and RSSI is the measured RSSI in dBm. \cite{5415423}
+We experimented with RSSI measurements to determine how well they worked by taking 2 Bluetooth modules, and measuring the RSSI while moving them apart.
+One remained stationary on the floor and the other was moved away from it 1 foot at a time.
+At each point, we took 3 RSSI measurements and averaged them. Figure \ref{fig:rssiplot} shows the results (the original data is in table \ref{table:rssidata}).
 
@@
 measured distances in feet, not meters, we could still use the above
-formula without worrying about unit conversions. The constants \(A\) and
-\(n\), provided we determined them empirically based on the data, would
+formula without worrying about unit conversions. The constants $A$ and
+$n$, provided we determined them empirically based on the data, would
 encode the conversions. As such, we fit the data using the above formula
-with \(A=-48\) and \(n=3\), resulting in the blue curve above. While the
+with $A=-48$ and $n=3$, resulting in the blue curve above. While the
 general shape of the curve matches, there is significant noise in the
@@
 \item
-  \emph{Communication between the two robots}. While Bluetooth may not
-  offer good distance measurement via RSSI, it can be used for reliable
-  communication between modules. It would be straightforward for one
-  hunting robot to inform the other whether it believes it is
-  approaching the beacon or not. In the simplest case, a hunting robot
-  that is approaching, or already at, the beacon can provide a second
-  point of reference for a currently hunting robot.
+  \emph{Communication between the two robots}. While Bluetooth may not offer good distance measurement via RSSI, it can be used for reliable  communication between modules.
+  It would be straightforward for one hunting robot to inform the other whether it believes it is approaching the beacon or not.
+  In the simplest case, a robot that is approaching, or already at, the beacon can provide a second point of reference for a currently hunting robot.
 \item
-  \emph{More complete usage of IMU}.
-Additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
-This, paired with inter-swarm
-  communication, would make for a much more sophisticated and likely
-  much more efficient system. (Of course, this does not resolve the
-  shallow gradients problem -- but it would allow the approach to the
-  beacon to be much faster.)
+  \emph{More complete usage of IMU}. Additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
+  This, paired with inter-swarm communication, would make for a much more sophisticated and likely much more efficient system.
+  (Of course, this does not resolve the shallow gradients problem -- but it would allow the approach to the beacon to be much faster.)
 \end{itemize}
 
-\subsubsection{Appendix A: Bill Of Materials}
+\bibliography{references}{}
+\bibliographystyle{ieeetr}
 
+\begin{appendices}
+  \section{Bill Of Materials}
   This project was done with a \$150 budget in mind.
@@
 
+  \begin{centering}
+  \tiny
   \begin{longtable}[]{@{}lllllll@{}}
@@
   \end{longtable}
+  \end{centering}
 
   \textbf{Total:} \$97.68
-
-\bibliography{references}{}
-\bibliographystyle{ieeetr}
+\end{appendices}
 

### 2018-02-06 article/article.tex
Basic restructuring + delisting

@@
 
+% TODO: motivation, RSSI, fake chips
+
 In this article, we investigate using the Received Signal Strength Indicator (RSSI) of Bluetooth Low Energy (BLE) 4.0 chips to allow wheeled mobile robots to navigate towards a stationary base station.
@@
 
-\subsection{Hardware}
+\subsection{Design}
 
 \subsubsection{Chassis}
+
+% Possible subsections?
+% - building robot
+% - signal detection
+% - navigation
 
@@
 
-\paragraph{Servos}
+% \paragraph{Servos}
 
@@
 
-\paragraph{Gradient descent}
+% \paragraph{Gradient descent}
 
@@
 There are many potential avenues for improvements or further
-development:
+development.
 
-\begin{itemize}
-\tightlist
-\item
-  \emph{Communication between the two robots}. While Bluetooth may not offer good distance measurement via RSSI, it can be used for reliable  communication between modules.
+Firstly, \emph{Communication between the two robots}: while Bluetooth may not offer good distance measurement via RSSI, it can be used for reliable  communication between modules.
 It would be straightforward for one hunting robot to inform the other whether it believes it is approaching the beacon or not.
 In the simplest case, a robot that is approaching, or already at, the beacon can provide a second point of reference for a currently hunting robot.
-\item
-  \emph{More complete usage of IMU}. Additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
+
+Also, \emph{More complete usage of IMU}: additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
 This, paired with inter-swarm communication, would make for a much more sophisticated and likely much more efficient system.
 (Of course, this does not resolve the shallow gradients problem -- but it would allow the approach to the beacon to be much faster.)
-\end{itemize}
 

### 2018-02-06 article/article.tex
Bruce fixes + refs

@@
 
-\subsubsection{Chassis}
-
-% Possible subsections?
-% - building robot
-% - signal detection
-% - navigation
+\subsubsection{Building the Robot}
 
@@
 
-\begin{figure}
-  \centering
-  \subfloat[Top view]{
-    \includegraphics[width=0.4\textwidth]{top_view.jpg}
-    \label{fig:robottop}
-  }
-  \subfloat[Side view]{
-    \includegraphics[width=0.4\textwidth]{side_view.jpg}
-    \label{fig:robotside}
-  }
-  \\
-  \subfloat[Front view]{
-    \includegraphics[width=0.4\textwidth]{front_view.jpg}
-    \label{fig:robotfront}
-  }
-  \subfloat[Back view]{
-    \includegraphics[width=0.4\textwidth]{inside_view.jpg}
-    \label{fig:robotback}
-  }
-  \caption{Multiple views of the completed robots}
-  \label{fig:robotviews}
-\end{figure}
-
-\hypertarget{electronics}{%
-\subsubsection{Electronics}\label{electronics}}
-
-Each robot consisted of 4 main components:
-\begin{enumerate}
-  \item PIC32MC250F128B microcontroller.
-  \item MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer.
-  \item HM-10 BLE module with a Texas Instruments CC2541 chip.
-  \item Two 9 gram FS90R micro continuous rotation servos.
-\end{enumerate}
+% TODO: move around to be more top-down
+Each robot consisted of 4 main components: a PIC32MC250F128B microcontroller, a MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer, a HM-10 BLE module with a Texas Instruments CC2541 chip, and two 9 gram FS90R micro continuous rotation servos.
 
@@
 
-\paragraph{Bluetooth modules}
+For the drive system, we used FS90R servos.
+These servos are continuous rotation servos with a stall torque of 1.5 kg-cm.
+Like most servos, they are driven with PWM signals at 50 hertz (20 millisecond periods).
+A duty cycle with about 1.5 milliseconds on is roughly neutral.
+With the PIC32s running at 40 MHz, a prescaler of 32 and a maximum count of 25600 produces the required 50 hertz signal ($25600/\left(40\cdot 10^6 \frac{1}{\text{s}}/ 32 \right) = 0.02048 \text{ ms}$).
+The PWM values (in timer counts) thus ranged from ranging from 1280 to 2560.
 
-The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao,
-\footnote{Jnhuamao is the company which makes the HM-10 module, along with other Bluetooth modules. Their English website is: \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}
-and did not come with genuine Jnhuamao firmware.
+The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
+compiler and the PIC 32 Legacy Peripheral Library (plib).
+The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}
+
+\subsubsection{Signal Detection: Bluetooth Modules}
+
+% Jnhuamao is the company which makes the HM-10 module, along with other Bluetooth modules. Their English website is: \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.
+
+The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao , and did not come with genuine Jnhuamao firmware.
 However, they did have genuine TI CC2541 chips.
 We realized they were fakes when they did not behave according to
-the Jnhuamao data sheet.
-\footnote{The original data sheet comes from \cite{jnhuamaodatasheet}.
-A PDF is available at \cite{jnhuamaomit}, which may be out of date, but is eaiser to get.}
+the Jnhuamao data sheet at \cite{jnhuamaodatasheet}, 
+with a PDF available at \cite{jnhuamaomit}, which may be out of date, but is easier to get.
 Luckily, the hardware on the fake chips is the same as that of the genuine chips, minus an external crystal, and the genuine firmware checks for the presence of the crystal, and works even without
@@
 
-\subsection{Software}
-
-The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
-compiler and the PIC 32 Legacy Peripheral Library (plib). \footnote{
-The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}}
-
-\subsubsection{BLE}
-
 The BLE module used UART, at 9600 baud 8N1. We used \texttt{UART1} on
@@
 We used the following commands to interact with the bluetooth chip:
+
+% TODO: Tablify
 
@@
 
-\paragraph{IMU}
+\subsubsection{Navigation}
 
-The PIC communicates with the IMU via I2C. The QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module has 2 dies:
-one contains the 3-axis gyroscope and 3-axis accelerometer, which were not
-used in this project, and the other die is the AK8963 3-axis
-magnetometer (compass). \cite{ak8963cdatasheet}
+After measuring the proximity to the beacon, the robots would need to move towards it.
+The robot made use of the magnetometer (compass) inside the IMU to orient itself, and the gradient descent algorithm to decide the direction of its next movement.
 
-The compass is connected to the rest of the MPU module via an auxiliary I2C bus.
-In order to make the compass an accessible slave on the I2C bus, pass-through mode to be enabled on the IMU.
-\footnote{We based our IMU code off of another ECE 4760 project, the Self-Balancing Robot. \cite{selfbalancingrobot} Their \texttt{i2c\_helper.h} was very helpful.}
+For the IMU, we used the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module.
+It has 2 dies: one is the AK8963 3-axis magnetometer \cite{ak8963cdatasheet}, and the other contains the 3-axis gyroscope and 3-axis accelerometer, which were not used in this project.
+
+The microcontroller communicates with the IMU via I2C, and the compass is connected to the rest of the MPU module via an auxiliary I2C bus.
+For communication between the microcontroller and the AK8963's main die, we based our work off of basic I2C functions from another ECE 4760 project, the Self-Balancing Robot \cite{selfbalancingrobot}.
+In particular, their \texttt{i2c\_helper.h} was very helpful.
+
+For communication with the actual magnetometer, we needed to configure the IMU to set the compass must as a slave on the I2C bus.
+For this, it was necessary to enable pass-through mode on the IMU during its configuration.
+
 The AK8963 has several modes of operation, and the chip must be set to power-down mode before switching to other modes.
-We read compass values with single measurement mode, as in figure \ref{fig:imu_single_measurement}:
+We read compass values with single measurement mode, as in figure \ref{fig:imu_single_measurement}.
 
-\begin{enumerate}
-\item
-  Set the compass to single measurement mode in 14 bit resolution.
-\item
-  Read the 6 data registers (X low, X high, Y low, Y high, Z low, Z
-  high)
-\item
-  Read the Status 2 register to check for magnetic sensor overflow.
-  Without reading this register, the read is not considered complete and
-  further reads will fail.
-\item
-  Wait; if the IMU is read too frequently, it will not have enough time
-  to take measurements.
-\end{enumerate}
+A single compass read involves setting the compass to single measurement mode in 14 bit resolution, reading the 6 data registers (X low, X high, Y low, Y high, Z low, Z high), reading the Status 2 register to check for magnetic sensor overflow (without reading this register, the read is not considered complete and further reads will fail), and finally waiting to ensure that the IMU is not read too frequently, in which case it will not have enough time to take measurements.
 
@@
 that data to scale and center the magnetometer readings.
-
-% \paragraph{Servos}
-
-For the drive system, we used FS90R servos.
-These servos are continuous rotation servos with a stall torque of 1.5 kg-cm.
-Like most servos, they are driven with PWM signals at 50 hertz (20 millisecond periods).
-A duty cycle with about 1.5 milliseconds on is roughly neutral.
-With the PIC32s running at 40 MHz, a prescaler of 32 and a maximum count of 25600 produces the required 50 hertz signal ($25600/\left(40\cdot 10^6 \frac{1}{\text{s}}/ 32 \right) = 0.02048 \text{ ms}$).
-The PWM values (in timer counts) thus ranged from ranging from 1280 to 2560.
-
-% \paragraph{Gradient descent}
 
@@
 determination using Bluetooth, a generally unconventional approach.
-We knew that Bluetooth RSSI would be noisy, mostly due to multipath interference.
+We knew that Bluetooth RSSI would be noisy, mostly due to multipath interference and the presence of multiple transmitters in testing environments.
 The robots worked reliably when they stayed within roughly 1 meter of the beacon.

### 2018-02-06 article/article.tex
add raw convergence table

@@
 \newcommand{\NormalTok}[1]{#1}
-\usepackage{longtable,booktabs}
+\usepackage{longtable,booktabs,multirow}
+\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
+
 % Fix footnotes in tables (requires footnote package)
@@
 
+\begin{table}[]
+\centering
+\ra{1.3}
+\caption{Convergence performance of the robots}
+\label{table:convergence}
+\begin{tabular}{@{}lllllll@{}}
+\toprule
+\multirow{2}{*}{Trial} & \multicolumn{3}{c}{Robot 0}       & \multicolumn{3}{c}{Robot 1}      \\ \cmidrule(l){2-4} \cmidrule(l){5-7}
+                       & Time (s)  & Distance (cm) & Start & Time (s) & Distance (cm) & Start \\ \midrule
+1                      & 213       & 53            & South & 94       & 22            & North \\
+2                      & \emph{Collision} & 420           & South & 153      & 36            & North \\
+3                      & 308       & 19            & South & 105      & 36            & North \\
+4                      & \emph{Collision} & 340           & South & 80       & 20            & North \\
+5                      & \emph{Collision} & 240           & South & 92       & 30            & North \\
+6                      & 179       & 46            & North & 117      & 20            & South \\
+7                      & 56        & 45            & North & 162      & 24            & South \\
+8                      & 59        & 27            & North & 197      & 26            & South \\
+9                      & 59        & 19            & North & 79       & 30            & South \\
+10                     & \emph{Collision} & 231           & North & 67       & 23            & South \\ \bottomrule
+\end{tabular}
+\end{table}
+
 While the chip reported RSSI in units proportional to dBm, and we

### 2018-02-06 article/article.tex
Delete bom

@@
 
-\begin{appendices}
-  \section{Bill Of Materials}
-  This project was done with a \$150 budget in mind.
-Costs for parts labeled ``Lab rental'' were borrowed from the department of Electrical and Computer Engineering at Cornell University, but can also be found at vendors like Digikey, Adafruit, and Amazon.
-
-  \begin{centering}
-  \tiny
-  \begin{longtable}[]{@{}lllllll@{}}
-  \toprule
-  \begin{minipage}[b]{0.15\columnwidth}\raggedright
-  Name\strut
-  \end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright
-  Manufacturer Part Number\strut
-  \end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
-  Vendor\strut
-  \end{minipage} & \begin{minipage}[b]{0.17\columnwidth}\raggedright
-  Vendor Part Number\strut
-  \end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
-  Quantity\strut
-  \end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright
-  Unit Cost\strut
-  \end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright
-  Total Cost\strut
-  \end{minipage}\tabularnewline
-  \midrule
-  \endhead
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  BLE 4.0 Module (TI CC2541)\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  HM-10\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  \href{https://www.ebay.com/itm/AT-09-BLE-Bluetooth-4-0-Uart-Transceiver-Module-CC2541-Central-Switching-HM-10/142425748901?ssPageName=STRK\%3AMEBIDX\%3AIT\&_trksid=p2057872.m2749.l2649}{Ebay}\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  142425748901\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  3\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$3.99\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$11.97\strut
-  \end{minipage}\tabularnewline
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  FEETECH FS90R (pack of 2) Continuous Rotation Robotic Servo\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  FS90R\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  \href{https://www.amazon.com/gp/product/B074BFQC3Q/ref=oh_aui_detailpage_o06_s00?ie=UTF8\&psc=1}{Amazon}\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  B074BFQC3Q\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  2\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$12.39\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$24.78\strut
-  \end{minipage}\tabularnewline
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  HiLetgo 9-Axis 9 DOF 16 Bit Gyroscope Acceleration Magnetic Sensor\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  MPU-9250\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  \href{https://www.amazon.com/HiLetgo-Gyroscope-Acceleration-Accelerator-Magnetometer/dp/B01I1J0Z7Y/ref=redir_mobile_desktop?_encoding=UTF8\&dpID=51nl2fcMh6L\&dpPl=1\&keywords=mpu\%209250\&pi=AC_SX236_SY340_QL65\&qid=1512564044\&ref=plSrch\&ref_=mp_s_a_1_3\&sr=8-3}{Amazon}\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  B01I1J0Z7Y\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  2\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$8.49\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$16.98\strut
-  \end{minipage}\tabularnewline
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  Small board\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  Lab rental\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  3\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$5.00\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$15.00\strut
-  \end{minipage}\tabularnewline
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  PIC32MX250F128B\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  PIC32MX250F128B\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  Lab rental\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  3\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$5.00\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$15.00\strut
-  \end{minipage}\tabularnewline
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  6-inch Protoboard\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  Lab rental\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  3\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$2.50\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$7.50\strut
-  \end{minipage}\tabularnewline
-  \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  Male header pins\strut
-  \end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-  Lab\strut
-  \end{minipage} & \begin{minipage}[t]{0.17\columnwidth}\raggedright
-  --\strut
-  \end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-  129\strut
-  \end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright
-  \$0.05\strut
-  \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
-  \$6.45\strut
-  \end{minipage}\tabularnewline
-  \bottomrule
-  \end{longtable}
-  \end{centering}
-
-  \textbf{Total:} \$97.68
-\end{appendices}
-
 \end{document}

### 2018-02-06 article/article.tex
add average convergence table
More footnotes removed

@@
 
-  The firmware file came from the Arduino forum post. \footnote{The firmware file can be found here:  \url{http://forum.arduino.cc/index.php?action=dlattach;topic=393655.0;attach=183702}}
+  The firmware file came from the Arduino forum post. \cite{firmwarefile}
 \end{enumerate}
@@
 
+\begin{table}[]
+\centering
+\ra{1.3}
+\caption{Average convergence behavior}
+\label{table:averagecon}
+\begin{tabular}{@{}lll@{}}
+\toprule
+                       & Robot 0        & Robot 1     \\ \midrule
+Convergence time (s)   & 146            & 115         \\
+Convergence rate       & 60 \%          & 100 \%      \\
+Final distance (cm)    & 35             & 27          \\
+Failure distance (cm)  & 308            & --          \\ \bottomrule
+\end{tabular}
+\end{table}
+
 While the chip reported RSSI in units proportional to dBm, and we

### 2018-02-06 article/article.tex
Collision -> timeout trial 2

@@
 1                      & 213       & 53            & South & 94       & 22            & North \\
-2                      & \emph{Collision} & 420           & South & 153      & 36            & North \\
+2                      & \emph{Timeout} & 420           & South & 153      & 36            & North \\
 3                      & 308       & 19            & South & 105      & 36            & North \\

### 2018-02-06 article/article.tex
Fix ref

@@
 \usepackage{subfig}
-\usepackage{float}
 \usepackage[page]{appendix}
@@
 
-% Jnhuamao is the company which makes the HM-10 module, along with other Bluetooth modules. Their English website is: \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.
-
-The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao , and did not come with genuine Jnhuamao firmware.
+The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao \cite{jnhuamao}, the manufacturer of the HM-10 module we intended to use.
+The modules we bought did not come with genuine Jnhuamao firmware.
 However, they did have genuine TI CC2541 chips.

### 2018-02-12 article/article.tex
start editing, about halfway there

@@
 
-\subsubsection{Building the Robot}
-
-The robots are made from 4 3D printed pieces: 2 wheels, the frame, and
-the caster in the back.
-Figure \ref{fig:robotchassis} shows the main components.
-The servos, a 3 AA battery holder, and a perfboard containing all the circuitry are mounted directly to the frame.
+The system consists of 2 robots, and a base station.
+We designed the robots to be small, light, and maneuverable.
+Each robot has 2 wheels, driven by a 9 gram micro servo, as well as a third round plastic leg, serving as a caster wheel.
+We 3D printed the robot's frame in ABS, resulting in an easy to assemble robot.
+After printing the parts, we mounted a 3 AA battery pack to the frame to provide power,
+and screwed in a perfboard on standoffs. See \ref{fig:robotchassis} for the parts in each robot.
 
@@
 
-The parts were printed in ABS using Maker Select 3D Printer v2 printers.
-All parts were printed with a layer height of 0.3 mm, as there was no need for a smooth finish or high tolerances.
-The parts were designed in OpenSCAD, and sliced with Cura.
-
-% TODO: move around to be more top-down
-Each robot consisted of 4 main components: a PIC32MC250F128B microcontroller, a MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer, a HM-10 BLE module with a Texas Instruments CC2541 chip, and two 9 gram FS90R micro continuous rotation servos.
-
-The PIC32 and supporting circuity was on a small PCB (the ``Small Dev Board'' from ECE 4760 at Cornell University, the class during which we built this) shown in figure \ref{fig:schematic-sdb}.
-The rest of the components were assembled on solderable breadboard as shown in figure \ref{fig:schematic-robot}.
+The electronics for each robot consisted of 4 main components: a PIC32MC250F128B microcontroller, a MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer, a HM-10 BLE module with a Texas Instruments CC2541 chip, and two 9 gram FS90R micro continuous rotation servos.
+The BLE module connected to the PIC through a UART connection, and the IMU connected to the PIC through an I2C bus.
+The base station was identical to the robots, minus the IMU and servos.
+The HM-10 Bluetooth module was mounted perpendicular to the breadboard, sticking up in the air to hopefully improve the performance of the radio.
+Figure \ref{fig:schematic} provides a full schematic for one of the robots.
 
@@
 
-For the drive system, we used FS90R servos.
-These servos are continuous rotation servos with a stall torque of 1.5 kg-cm.
-Like most servos, they are driven with PWM signals at 50 hertz (20 millisecond periods).
-A duty cycle with about 1.5 milliseconds on is roughly neutral.
-With the PIC32s running at 40 MHz, a prescaler of 32 and a maximum count of 25600 produces the required 50 hertz signal ($25600/\left(40\cdot 10^6 \frac{1}{\text{s}}/ 32 \right) = 0.02048 \text{ ms}$).
-The PWM values (in timer counts) thus ranged from ranging from 1280 to 2560.
+While trying to use the Bluetooth modules (which we bought off Ebay), we discovered that they didn't adhere to the data sheet \cite{jnhuamaodatasheet}, and after doing some research, concluded they were fake!
+While the boards still had TI CC2541 chips, the firmware they were running was not the genuine Jnhuamao firmware.
+Luckily, the only difference between the fake boards and the genuine boards was that then genuine boards had an external crystal, but the genuine firmware checks for the presence of the crystal, and works even without it.
+\cite{crystal}
+Because of that, we were able to salvage the fake chips by reprogramming them with genuine firmware according to an Arduino Fourm post \cite{crystal}.
+Essentially, we connected the fake chips to an Arduino Teensy (which is 3.3 volts, and so won't damage the 3.3 volt CC2541) as indicated in table \ref{table:progpins}, using the schematic in figure \ref{fig:hm10} for reference.
 
-The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
-compiler and the PIC 32 Legacy Peripheral Library (plib).
-The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}
-
-\subsubsection{Signal Detection: Bluetooth Modules}
-
-The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao \cite{jnhuamao}, the manufacturer of the HM-10 module we intended to use.
-The modules we bought did not come with genuine Jnhuamao firmware.
-However, they did have genuine TI CC2541 chips.
-We realized they were fakes when they did not behave according to
-the Jnhuamao data sheet at \cite{jnhuamaodatasheet},
-with a PDF available at \cite{jnhuamaomit}, which may be out of date, but is easier to get.
-Luckily, the hardware on the fake chips is the same as that of the genuine chips, minus an external crystal, and the genuine firmware checks for the presence of the crystal, and works even without
-it. \cite{crystal}
-As such, we reprogrammed the chips with the genuine firmware according
-to an Arduino Fourm post \cite{crystal}:
-
-\begin{enumerate}
-\item
-  We soldered wires to the programming pins on the breakout boards, and connected those pins to an Arduino Teensy 3.2.
-  We chose a Teensy because it is 3.3 volts as opposed to 5, which would damage the 3.3 volt CC2541. The pins were connected as follows:
-
-  \begin{longtable}[]{@{}lll@{}}
+\begin{table}[]
+  \centering
+  \begin{tabular}{@{}lll@{}}
     \toprule
@@
     \midrule
-  \endhead
     \texttt{DEBUG\_CLOCK} & 7 & 5\tabularnewline
@@
     \bottomrule
-  \end{longtable}
+  \end{tabular}
+  \caption{Arduino to CC2541 pin connections for programming.}
+  \label{table:progpins}
+\end{table}
 
-  Figure \ref{fig:hm10} shows the layout of the HM-10 module.
 \begin{figure}
@@
 
-\item
-  We uploaded the \texttt{CCLoader.ino} \cite{ccloader} sketch to the Arduino.
-\item
-  Finally, we ran (in a Windows virtual machine, due to the dubious origin of the software) \texttt{CCLoader.exe} \cite{ccloaderexe}.
-  This program takes 3 arguments:
-
-  \begin{Verbatim}[gobble=4]
+Then, we uploaded the \texttt{CCLoader.ino} \cite{ccloader} sketch to the Arduino.
+This sketch bit bangs the programming signals for the CC2541/
+Finally, we ran (in a Windows virtual machine, due to the dubious origin of the software) \texttt{CCLoader.exe} \cite{ccloaderexe}:
+\begin{Verbatim}[gobble=2]
   CCLoader.exe <COM Port> <Firmware.bin> 0
 \end{Verbatim}
+The firmware file came from the Arduino forum post. \cite{firmwarefile}
 
-  The firmware file came from the Arduino forum post. \cite{firmwarefile}
-\end{enumerate}
+Next, we updated the chips firmware from from 540, the version we just flashed onto it, to the latest (at the time we did this project) released version, 603. \cite{jnhuamao603}
+We connected the HM-10 module to a computer using a 3.3 V FTDI to USB adapter.
+Then, using PUTTY established a serial connection (9600 baud, 8N1) and sent the chip \texttt{AT}. If it is connected properly, it will respond with \texttt{OK}.
+Next, send the chip \texttt{AT+SBLUP} to put it in firmware update mode.
+It will respond with \texttt{OK+SBLUP}. Terminate the PUTTY session.
+Run the \texttt{HMSoft.exe} program distributed in the firmware update download.
+(Again in a Windows virtual machine, due to the dubious origin of the software.)
+Select the proper port and firmware file using the software, and hit ``Load Image''.
+The software should handle the rest!
+To make sure it worked, establish a serial connection again using PUTTY.
+Send \texttt{AT+VERS?} to query the chip for version information.
 
-After flashing genuine firmware onto the chip, the next step was to update the firmware to the latest version.
-The firmware flashed onto the board was version 540, but Jnhuamao had (at the time we did this project) released version 603. \cite{jnhuamao603}
-Based on their instructions \cite{jnhuamaoinstructions}, we flashed our chips:
+% \cite{jnhuamaoinstructions}
 
-\begin{enumerate}
-\item
-  Connect the HM-10 module to a computer using a 3.3 V FTDI to USB
-  adapter. Then, use PUTTY to establish a serial session (9600 baud,
-  8N1). Send the chip \texttt{AT}; if it is connected properly, it will
-  respond with \texttt{OK}.
-\item
-  Send the chip \texttt{AT+SBLUP} to put it in firmware update mode. It
-  will respond with \texttt{OK+SBLUP}. Terminate the PUTTY session.
-\item
-  Run the \texttt{HMSoft.exe} program distributed in the firmware update
-  download. (Again in a Windows virtual machine, due to the dubious origin of the software.)
-\item
-  Select the proper port and firmware file using the software, and hit
-  ``Load Image''. The software should handle the rest!
-\item
-  To make sure it worked, establish a serial connection again using
-  PUTTY. Send \texttt{AT+VERS?} to query the chip for version
-  information.
-\end{enumerate}
-
-The BLE module used UART, at 9600 baud 8N1. We used \texttt{UART1} on
-the PIC for communicating with the BLE module, and used \texttt{UART2}
-for communicating with the computer (for debugging).
-We used the following commands to interact with the bluetooth chip:
-
-% TODO: Tablify
+At this point, we finally had the bluetooth chips in a working state.
 

### 2018-02-15 article/article.tex
finish first pass of refactoring

@@
 \usepackage{longtable,booktabs,multirow}
+\usepackage{tabularx}
 \newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
@@
 At this point, we finally had the bluetooth chips in a working state.
+In order to have them measure signal strength, we sent it the sequence of commands
+shown in table \ref{table:blecommands} in order to obtain a signal strength measurement.
+One interesting thing to note about the chip is that commands do not have to end with newlines or carriage returns.
+However, if sent, the chip will ignore them.
 
-\begin{itemize}
-\item
-  \texttt{AT+RESET}: resets the chip to ensure it is in a clean state
-  before receiving other commands.
-\item
-  \texttt{AT+IBEA1}: enables the iBeacon functionality of the chip (sets
-  it to 1; \texttt{AT+IBEA0} would disable it by setting it to 0). This
-  allows the chip to be found with an RSSI scan. After setting the
-  value, we query it with \texttt{AT+IBEA?} and send the result over the
-  other serial port to a computer for debugging.
-\item
-  \texttt{AT+ROLE0} or \texttt{AT+ROLE1}: sets the role to either
-  peripheral (0) or central (1). The base station is set to peripheral,
-  and the 2 robots are set to central. Peripheral means the device will
-  respond in inquiries from a central device. This allows it to be
-  discovered during an RSSI scan. After setting the value, we query it
-  with \texttt{AT+ROLE?} and send the result over the other serial port
-  to a computer for debugging.
-\item
-  \texttt{AT+IMME0} or \texttt{AT+IMME1}: sets the work state of the
-  device to either actively listening for Bluetooth signals (0), or only
-  acting when it receives a serial command (1). Once again, the base
-  station is set to 0: it needs to listen for signals and respond. The
-  robots are set to 1, as the chips need to initiate scan requests when
-  they receive the command over serial. After setting the value, we
-  query it with \texttt{AT+IMME?} and send the result over the other
-  serial port to a computer for debugging.
-\item
-  \texttt{AT+NAME\%s}: sets the name of the chip (which is visible when
-  scanning) to \texttt{\%s} (For example, \texttt{AT+NAMEPIRATE} names
+\begin{table}[]
+\centering
+\ra{1.3}
+\caption{HM-10 Bluetooth module signal strength measurement commands}
+\label{table:blecommands}
+\begin{tabularx}{\textwidth}{@{}l X@{}}
+\toprule
+Command & Description \\ \midrule
+\texttt{AT+RESET} & Resets the chip to ensure it is in a clean state
+before receiving other commands. \\
+
+\texttt{AT+IBEA1} & Sets the iBeacon functionality of the chip (1 on, 0 off).
+This allows the chip to be found with an RSSI scan. \\
+
+\texttt{AT+ROLE\{0|1\}} & Sets the role to either peripheral (0) or central (1).
+The base station is set to peripheral, and the 2 robots are set to central.
+Peripheral means the device will respond in inquiries from a central device.
+This allows it to be discovered during an RSSI scan.\\
+
+\texttt{AT+IMME\{0|1\}} & Sets the work state of the device to either actively
+listening for Bluetooth signals (0), or only acting when it receives a serial
+command (1). Once again, the base station is set to 0: it needs to listen for
+signals and respond. The robots are set to 1, as the chips need to initiate scan
+requests when they receive the command over serial. \\
+
+\texttt{AT+NAME\{str\}} & Sets the name of the chip (which is visible when
+scanning) to the string \texttt{str} (For example, \texttt{AT+NAMEPIRATE} names
 the chip \texttt{PIRATE}). We give all the chips unique names to make
-  debugging easier. After setting the value, we query it with
-  \texttt{AT+NAME?} and send the result over the other serial port to a
-  computer for debugging.
-\item
-  \texttt{AT+SHOW3}: configures the device to advertise both its name
-  and RSSI when scanning.
-\item
-  \texttt{AT+ADDR?}: queries the device for its hardware address. We
+debugging easier. \\
+
+\texttt{AT+SHOW3} & Configures the device to advertise both its name
+and RSSI when scanning. \\
+
+\texttt{AT+ADDR?} & Queries the device for its hardware address. We
 recorded the hardware device of each chip, as when doing RSSI scans,
-  the results are reported by hardware address.
-\item
-  \texttt{AT+DISI?}: performed only on the robots, causing a discovery
+the results are reported by hardware address. \\
+
+\texttt{AT+DISI?} & Performed only on the robots, causing a discovery
 scan. The result of the scan is a series of lines of the form:
-
-\begin{verbatim}
-OK+DISC:00000000:00000000000000000000000000000000:
-  0000000000:6832A3801EBE:-080
-\end{verbatim}
-
+{\begin{align*}
+&\mathtt{OK+DISC:00000000:00000000000000000000000000000000:}\\
+&\mathtt{  0000000000:6832A3801EBE:-080}
+\end{align*}}
 The second to last token, \texttt{6832A3801EBE}, is the hardware
@@
 finds (``line'' is a misnomer as it does not separate them with any
-  characters), followed by \texttt{OK+DISCE}.
-\end{itemize}
+characters), followed by \texttt{OK+DISCE}. \\
+\bottomrule
+\end{tabularx}
+\end{table}
 
-One interesting thing to note about the chip is that commands do not have to end with newlines or carriage returns.
-However, if sent, the chip will ignore them.
-
-\subsubsection{Navigation}
-
-After measuring the proximity to the beacon, the robots would need to move towards it.
-The robot made use of the magnetometer (compass) inside the IMU to orient itself, and the gradient descent algorithm to decide the direction of its next movement.
-
-For the IMU, we used the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module.
-It has 2 dies: one is the AK8963 3-axis magnetometer \cite{ak8963cdatasheet}, and the other contains the 3-axis gyroscope and 3-axis accelerometer, which were not used in this project.
+Once we had the Bluetooth signal strength working, we configured the IMU,
+a QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module.
+It has 2 dies: one is the AK8963 3-axis magnetometer \cite{ak8963cdatasheet},
+and the other contains the 3-axis gyroscope and 3-axis accelerometer, which we did not use.
 
@@
 In order to calibrate the compass, the robots spun in place when powered
-on.
-They recorded the maximum and minimum values for each axes, and used
+on. They recorded the maximum and minimum values for each axes, and used
 that data to scale and center the magnetometer readings.
 
-The algorithm for deciding what path to follow is a basic version of
-gradient descent. Figure \ref{fig:graddesc} shows 2 versions of the gradient descent algorithm.
+\subsubsection{Navigation}
+
+Once we had the robots working, we turned to the algorithm.
+Since they don't have much data -- only the last couple signal strength measurements --
+we decided to use a simple gradient descent algorithm.
+Figure \ref{fig:graddesc} shows 2 versions of the gradient descent algorithm.
 
@@
 
-In most cases, at least 1 of the 2 robots successfully made it to the
-base station. However, it was not as reliable as we initially hoped. One
-of the main reasons for this was the noise in RSSI measurements.
+On average, when starting from 1 meter away, the robots successfully made it to the base station 80\% of the time,
+within an average time of 126 seconds.
+This wasn't as reliable, or as fast, as we initially hoped.
+One of the main reasons for this was the noise in RSSI measurements.
 
@@
 
-Despite how noisy the RSSI measurements where, the robots were still
-able to perform a reasonably accurate gradient descent. In most cases,
-at least one of the 2 robots would find the beacon in a matter of
-minutes.
+Despite the noise in the RSSI readings, the robots did surprisingly well.
+We tested the robots by starting the both 1 meter from the base station,
+one on the north side, and one on the south side.
+After they started, we timed the amount of time it took them to arrive
+at the base station and flash their LEDs.
+We set a timeout of 5 minutes, at which point we would measure the
+distance between the robot and the base station.
+We also measured the distance when the robot arrived.
+All distances were measured between the Bluetooth modules.
+The results of this are displayed in table \ref{table:convergence},
+with averages in table \ref{table:averagecon}.
+
+There are some interesting patterns in the data.
+One obvious one is that robot 1 always made it,
+while robot 0 only had a 60\% success rate.
+In addition, when robot 1 made it, it was on average 8 centimeters
+closer than robot 0, and converged about 30 seconds faster!
+We think that perhaps the antennas on each had different sensitivities,
+as both used the same RSSI threshold for stopping once they reached the goal.
+Another interesting result is that when robot 0 failed, it was
+on average 3 meters away -- far off in the land of shallow gradients.
+The only way to recover from that was dumb luck, as the signal would have been
+dominated by noise that far away.
 

### 2018-02-15 article/article.tex
fluff

--- 2018-02-06 article/article.tex
@@
 
-% TODO: motivation, RSSI, fake chips
+Navigating through a plane is a problem with many diverse applications and solutions.
+We explore the approach of letting a robot use the strength of a broadcasted signal to locate itself.
+In particular, we investigate using the Received Signal Strength Indicator (RSSI) of Bluetooth Low Energy (BLE) 4.0 chips to allow wheeled mobile robots to navigate towards a stationary base station.
+The robot detects its proximity to the station based on the strength of the signal and moves towards what it believes to be the signal source.
 
-In this article, we investigate using the Received Signal Strength Indicator (RSSI) of Bluetooth Low Energy (BLE) 4.0 chips to allow wheeled mobile robots to navigate towards a stationary base station.
+In the process of designing and building the hunting robots and the beacon, we encountered an interesting challenge in particular: fake Bluetooth modules from non-official vendors with usable hardware but out-of-date (and poorly documented) firmware.
+We explain the process we used to find, flash, and verify the correct firmware in order to use it for our project.
+
 Each robot was controlled by a PIC32MX250 microcontroller, and used a 3 axis magnetometer as as compass in order to reliably turn, as well as 2 micro 9 g servos to drive.
@@
 
-\subsubsection{Building the Robot}
+The design of the robots and beacon can be split into 3 rough categories: first, the construction of the robot itself, including the design of the chassis; second, the integration of the bluetooth modules that were used to transmit and receive signals, and lastly, the algorithm and sensors used by the robot to orient itself and move towards the source of the signal.
 
-The robots are made from 4 3D printed pieces: 2 wheels, the frame, and
+\subsubsection{Robot Construction}
+
+The robots themselves are made from 4 3D printed pieces: 2 wheels, the frame, and
 the caster in the back.
@@
 
-Firstly, \emph{Communication between the two robots}: while Bluetooth may not offer good distance measurement via RSSI, it can be used for reliable  communication between modules.
+Firstly, \emph{communication between the two robots}: while Bluetooth may not offer good distance measurement via RSSI, it can be used for reliable  communication between modules.
 It would be straightforward for one hunting robot to inform the other whether it believes it is approaching the beacon or not.
@@
 
-Also, \emph{More complete usage of IMU}: additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
+Also, \emph{more complete usage of IMU}: additional usage of the accelerometer and gyroscope, coupled with feedback from the servos, would allow the robots to maintain a dead-reckoning position estimate.
 This, paired with inter-swarm communication, would make for a much more sophisticated and likely much more efficient system.

### 2018-02-15 article/article.tex

--- 2018-02-15 article/article.tex
@@
 \subsection{Design}
+
+The design of the robots and beacon can be split into 3 rough categories: first, the construction of the robot itself, including the design of the chassis; second, the integration of the bluetooth modules that were used to transmit and receive signals, and lastly, the algorithm and sensors used by the robot to orient itself and move towards the source of the signal.
+
+\subsubsection{Robot Construction}
 
